---
title: "WorMotel lifespan Heritability"
author: "Tim C."
date: "5/17/2018"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---
This code uses processed survival data from Chris Fang-Yen's lab to calcualte heritability of survival traits. The data were collected with WorMotel in spring 2018 by Matt Churgin at UPenn. Two independent bleaches of 12 strains were run on three replicate wormmotels (20 nematodes / strain). Heritability is calculated 

```{r, echo=F, warning = F, message=F}
library(rio)
library(ggplot2)
library(dplyr)
library(survival)
library(flexsurv)
library(broom)
library(minpack.lm)
library(forcats)
library(glue)
library(scales)
library(boot)
try(setwd(dirname(rstudioapi::getActiveDocumentContext()$path)))
```

## Read in data and define functions
```{r, echo=FALSE, warning=FALSE}
dat <- as.data.frame(read.csv(file = "WildIsolateLSData_180515.csv", header = T)) %>%
  dplyr::mutate(lifespan_days = ifelse(lifespan_days == "NaN", NA, lifespan_days))

glimpse(dat)

# define heritability test
# data is data frame that contains strain and phenotype column
# indicies are used by the boot function to sample from the 'data' data.frame
H2.test.boot <- function(data, indicies){

  d <- data[indicies,]

  pheno <- as.data.frame(dplyr::select(d, phenotype))[,1]
  strain <- as.factor(d$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}
# data is data frame that contains strain and phenotype column
H2.test <- function(data){

  pheno <- as.data.frame(dplyr::select(data, phenotype))[,1]
  strain <- as.factor(data$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}

# df is data frame that contains strain and phenotype column
H2.calc <- function(df, boot = T){


  if(boot == T){
    # bootstrapping with 1000 replications
    results <- boot(data=df, statistic=H2.test.boot, R=1000)

    # get 95% confidence interval
    ci <- boot.ci(results, type="bca")

    H2_errors <- data.frame(H2 = ci$t0, ci_l = ci$bca[4], ci_r = ci$bca[5])

    return(H2_errors)

  } else {

    H2 <- data.frame(H2 = H2.test(data = df), ci_l = NA, ci_r = NA)
    return(H2)
  }

}


# remove outliers function
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}
```

### Survival plot
```{r, echo=FALSE, warning=FALSE}
# data shaping
dat_proc <- dat %>%
  dplyr::filter(!is.na(lifespan_days), activity_type == -1) %>%
  dplyr::group_by(replicate_num, plate_num, strain, worm_num) %>%
  dplyr::mutate(alive = ifelse(time_days >= lifespan_days, 0, 1)) %>%
  dplyr::group_by(replicate_num, plate_num, strain, time_days) %>%
  dplyr::summarise(frac_alive = sum(alive)/n(),
                   n = n())

# plot survival curves for all strains
raw_plot_all <- ggplot(dat_proc) +
  aes(x=time_days, y=frac_alive, color=strain) +
  geom_line() +
  theme_light() +
  labs(x="Day", y="Fraction alive") +
  theme(legend.position = "right") +
  facet_grid(replicate_num ~ plate_num, labeller = label_both)
raw_plot_all


```
<br> Survival curves look as expected. Replicate two appears to have less variable lifespans.

## Lifespan Heritability
```{r, echo=FALSE, warning=FALSE}
# lifespan dataset
df_h2 <- dat %>%
  dplyr::distinct(strain, replicate_num, plate_num, worm_num, lifespan_days) %>%
  dplyr::select(replicate_num, plate_num, strain, worm_num, lifespan_days) %>%
  dplyr::mutate(status = ifelse(is.na(lifespan_days), 0, 1)) %>%
  dplyr::mutate(plate_num = factor(plate_num)) 

# plot lifespan
lifespan_plot <- ggplot(df_h2) +
  geom_boxplot(aes(x = strain, y = lifespan_days), outlier.shape = NA) +
  geom_jitter(aes(x = strain, y = lifespan_days, color = as.factor(replicate_num)), width = .25, alpha = 0.5) +
  theme_light() +
  labs(x="", y="Lifespan (d)", colour = "replicate", title = "" ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
#lifespan_plot

# calculate heritability for lifespan
df_h2_cal <- df_h2 %>%
  dplyr::filter(!is.na(lifespan_days)) %>%
  dplyr::group_by(plate_num, strain) %>%
  dplyr::mutate(phenotype = mean(lifespan_days)) %>%
  dplyr::distinct(plate_num, strain, phenotype) %>%
  dplyr::mutate(trait = "lifespan") 
# finish heritability calculation
df_h2_only <- df_h2_cal %>%
  dplyr::ungroup() %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)

# plot means for plates
lifespan_plot <- ggplot(df_h2_cal) +
  geom_boxplot(aes(x = fct_reorder(strain, phenotype), y = phenotype), outlier.shape = NA) +
  geom_jitter(aes(x = strain, y = phenotype, color = as.factor(plate_num)), width = .25, alpha = 0.5) +
  theme_light() +
  labs(x="", y="Mean Lifespan (d)", colour = "plate", title = "Heritability: 0.03" ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
#lifespan_plot

# calculate heritability for two independent replicates 2 reps
df_h2_cal_ir <- df_h2 %>%
  dplyr::filter(!is.na(lifespan_days)) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(phenotype = mean(lifespan_days)) %>%
  dplyr::group_by(replicate_num, strain) %>%
  dplyr::mutate(phenotype = mean(lifespan_days)) %>%
  dplyr::distinct(replicate_num, strain, phenotype) %>%
  dplyr::mutate(trait = "lifespan")
# finish heritability calculation 2 reps
df_h2_only_ir <- df_h2_cal_ir %>%
  dplyr::ungroup() %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)


# plot means for replicates (avg for individual nematodes per plate avgeraged across all three plates in replicate)
lifespan_plot_ir <- ggplot(df_h2_cal_ir) +
  geom_boxplot(aes(x = fct_reorder(strain, phenotype), y = phenotype), outlier.shape = NA) +
  geom_jitter(aes(x = strain, y = phenotype, color = as.factor(replicate_num)), width = .25, alpha = 0.5) +
  theme_light() +
  labs(x="", y="Mean Lifespan (d)", colour = "replicate", title = "Lifespan Heritability: 3.021841e-17" ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
lifespan_plot_ir

# measure block effect
ls_block_effect <- lm(phenotype ~ replicate_num, data = df_h2 %>%
  dplyr::filter(!is.na(lifespan_days)) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(phenotype = mean(lifespan_days)) %>%
  dplyr::group_by(replicate_num, strain) %>%
  dplyr::mutate(phenotype = mean(lifespan_days)) %>%
  dplyr::distinct(replicate_num, strain, phenotype) %>%
  dplyr::mutate(trait = "lifespan") %>%
  dplyr::ungroup())
#summary(ls_block_effect)

# regress out block effect 2 reps
df_h2_cal_ir_regressed <- df_h2 %>%
  dplyr::filter(!is.na(lifespan_days)) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(phenotype = mean(lifespan_days)) %>%
  dplyr::group_by(replicate_num, strain) %>%
  dplyr::mutate(phenotype = mean(lifespan_days)) %>%
  dplyr::distinct(replicate_num, strain, phenotype) %>%
  dplyr::mutate(trait = "lifespan") %>%
  dplyr::ungroup() %>%
  dplyr::mutate(phenotype = residuals(lm(phenotype ~ replicate_num))) 

# finish heritability calculation 2 reps
df_h2_only_ir_regressed <- df_h2_cal_ir_regressed %>%
  dplyr::ungroup() %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)

# plot means for replicates (avg for individual nematodes per plate avgeraged across all three plates in replicate) with replicate regressed 2 reps
lifespan_plot_ir_reg <- ggplot(df_h2_cal_ir_regressed) +
  geom_boxplot(aes(x = fct_reorder(strain, phenotype), y = phenotype), outlier.shape = NA) +
  geom_jitter(aes(x = strain, y = phenotype, color = as.factor(replicate_num)), width = .25, alpha = 0.5) +
  theme_light() +
  labs(x="", y="Residual mean lifespan", colour = "replicate", title = "Assay regressed lifespan heritability: 0.2704377" ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
lifespan_plot_ir_reg
```
<br>There is a block effect, correcting for this effect increases lifespan heritability to 0.27.

## Fit weibull survival model
```{r, echo=FALSE, warning=FALSE}
# reshape data for survival object, ignore technical replicate and include just plate reps
dat_w <- dat %>%
  dplyr::distinct(strain, plate_num, worm_num, lifespan_days) %>%
  dplyr::select(strain, plate_num, worm_num, lifespan_days) %>%
  dplyr::mutate(status = ifelse(is.na(lifespan_days), 0, 1)) %>%
  dplyr::mutate(plate_num = factor(plate_num))

#Asign functions to calculate 90, 50, 10 percent survival times with Weibull model.
weibull_50 <- function(shape, scale) {
  qweibull(0.5, shape = shape, scale = scale)
}
weibull_10 <- function(shape, scale) {
  qweibull(0.9, shape = shape, scale = scale)
}
weibull_90 <- function(shape, scale) {
  qweibull(0.1, shape = shape, scale = scale)
}

# establish lists to populate with for loop
block_list <- list()
strain_list <- list()

# Calculate weibull shape and scale parameters for all strains in each plate
for(b in 1:length(unique(dat_w$plate_num))){
  
  block_df <- dat_w %>%
    dplyr::filter(plate_num == (unique(dat_w$plate_num)[b]))

  for(s in 1:length(unique(block_df$strain))){
  
    strain_df <- block_df %>%
      dplyr::filter(strain == (unique(block_df$strain)[s]))
    
    if(sum(strain_df$status) != 0){
    #run weibull fit and save parameters
    w_fit = flexsurvreg(Surv(lifespan_days, status) ~ 1, dist = "weibull", data = strain_df)
  
    w_90 <- summary(w_fit, fn = weibull_90, t=1)
  
    w_50 <- summary(w_fit, fn = weibull_50, t=1)
  
    w_10 <- summary(w_fit, fn = weibull_10, t=1)
  
    strain_surv <- data.frame(plate_num = unique(block_df$plate_num),
                  strain = unique(block_df$strain)[s],
                  shape = w_fit[["res"]][1],
                  scale = w_fit[["res"]][2],
                  LT10 = w_90[[1]][2],
                  LT50 = w_50[[1]][2],
                  LT90 = w_10[[1]][2])
    strain_surv <- strain_surv %>%
      dplyr::rename('LT10' = est, 'LT50' = est.1, 'LT90' = est.2)

    strain_list[[s]] <- strain_surv 

    } else {
      #fill in strain surv with NAs
      strain_surv <- data.frame(plate_num = unique(block_df$plate_num),
                  strain = unique(block_df$strain)[s],
                  shape = NA,
                  scale = NA,
                  LT10 = NA,
                  LT50 = NA,
                  LT90 = NA)

    strain_list[[s]] <- strain_surv 
  }
  }
  
  strain_df <- bind_rows(strain_list)
  
  block_list[[b]] <- strain_df
}

#combine all data
h_df <- dplyr::bind_rows(block_list) %>%
  dplyr::arrange(strain, plate_num)

# take average of model fits for three plates in each independent replicate
h_df_ir <- h_df %>%
  dplyr::mutate(replicate_num = ifelse(as.numeric(as.character(plate_num)) <= 3, 1, 2)) %>%
  dplyr::group_by(replicate_num, strain) %>%
  dplyr::mutate(shape = mean(shape, na.rm =T),
                scale = mean(scale, na.rm = T),
                LT10 = mean(LT10, na.rm = T),
                LT50 = mean(LT50, na.rm = T),
                LT90 = mean(LT90, na.rm = T)) %>%
  dplyr::distinct(replicate_num, strain, .keep_all = T)

# gather data for plot 6 reps
h_df_gather <- h_df %>%
  tidyr::gather(trait, phenotype, -plate_num, -strain)

# gather data for plot 2 reps
h_df_ir_gather <- h_df_ir %>%
  tidyr::gather(trait, phenotype, -replicate_num, -plate_num, -strain) %>%
  dplyr::select(-plate_num)

# plot parameters 6 reps
weibull_par_plot_raw <- ggplot(h_df_gather) +
  geom_boxplot(aes(x=strain, y=phenotype), color = "grey", outlier.shape = NA) +
  geom_jitter(aes(x=strain, y=phenotype, color = plate_num), width = 0.25, height = 0.1) +
  labs(x="", y="Weibull Parameter Values", color = "plate", title = "") +
  theme(legend.position = "right") +
  facet_grid(trait ~ ., scales = "free") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
#weibull_par_plot_raw

# plot parameters 2 reps
weibull_par_plot_raw <- ggplot(h_df_ir_gather) +
  geom_boxplot(aes(x=strain, y=phenotype), color = "grey", outlier.shape = NA) +
  geom_jitter(aes(x=strain, y=phenotype, color = as.factor(replicate_num)), width = 0.25, height = 0.1) +
  labs(x="", y="Weibull Parameter Values", color = "replicate", title = "") +
  theme(legend.position = "right") +
  facet_grid(trait ~ ., scales = "free") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
weibull_par_plot_raw

```
<br> N2 seems like an outlier for shape parameter.

## Calculate heritability of weibull fit parameters
```{r, echo=FALSE, warning=FALSE}
# calculate heritability for weibull fit parameters 6 reps
dat_wf_h <- h_df_gather %>%
  dplyr::ungroup() %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)

# calculate heritability for weibull fit parameters 2 reps
dat_wf_h_ir <- h_df_ir_gather %>%
  dplyr::ungroup() %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)

#Repeat heritability calculation without N2 6 reps
dat_wf_h_no_n2 <- h_df_gather %>%
  dplyr::filter(strain != "N2") %>%
  dplyr::ungroup() %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

#Repeat heritability calculation without N2 2 reps
dat_wf_h_ir_no_n2 <- h_df_ir_gather %>%
  dplyr::filter(strain != "N2") %>%
  dplyr::ungroup() %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

#heritability plot with and without N2 vs with N2, also plotting weibull fits and survival curves 6 reps
h2_6reps <- ggplot() +
  geom_point(data = dat_wf_h, aes(x = trait, y = H2), alpha = 0.5, color = "red", size = 4) +
  theme_light() +
  geom_jitter(data = dat_wf_h_no_n2, aes(x = trait, y = H2), alpha = 0.5, color = "black", size = 4, width= 0.1) +
  labs(x="Weibull fit parameter", y="Heritability", title = "6 reps: red = with N2, black = without N2")
#h2_6reps

weibull_par_plot_raw <- ggplot(h_df_gather) +
  geom_boxplot(aes(x=strain, y=phenotype), color = "grey") +
  geom_jitter(aes(x=strain, y=phenotype, color = plate_num), width = 0.1, height = 0.1) +
  labs(x="", y="Weibull Parameter Values", color = "plate") +
  theme_light() +
  theme(legend.position = "right") +
  facet_grid(trait ~ ., scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
#weibull_par_plot_raw

#heritability plot with and without N2 vs with N2, also plotting weibull fits and survival curves 2 reps
weibull_par_plot_raw <- ggplot(h_df_ir_gather) +
  geom_boxplot(aes(x=strain, y=phenotype), color = "grey") +
  geom_jitter(aes(x=strain, y=phenotype, color = as.factor(replicate_num)), width = 0.1, height = 0.1) +
  labs(x="", y="Weibull Parameter Values", color = "replicate") +
  theme_light() +
  theme(legend.position = "right") +
  facet_grid(trait ~ ., scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
#weibull_par_plot_raw

h2_2reps <- ggplot() +
  geom_point(data = dat_wf_h_ir, aes(x = trait, y = H2), alpha = 0.5, color = "red", size = 4) +
  theme_light() +
  geom_jitter(data = dat_wf_h_no_n2, aes(x = trait, y = H2), alpha = 0.5, color = "black", size = 4, width= 0.1) +
  labs(x="Weibull fit parameter", y="Heritability", title = "Weibull paramter heritability: red = with N2, black = without N2")
h2_2reps
```
<br> Heritability for shape parameter is high only because of N2 strain

## Weibull fitting examples
```{r, echo=FALSE, warning=FALSE}
# visualizing N2 fit vs. MY16 fits: Shape parameter high for N2 b/c variation in survival is low 6 reps
n2_surv_fit <- ggplot(dat_proc %>% dplyr::filter(strain =="N2" | strain == "MY16")) +
  aes(x=time_days, y=frac_alive, color = strain) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "right") +
  stat_function(fun=function(x)exp(1)^-((x/10.770605)^6.429521), geom="line", color = "#00BFC4") +
  stat_function(fun=function(x)exp(1)^-((x/11.217003)^6.524106), geom="line", color = "#00BFC4") +
  stat_function(fun=function(x)exp(1)^-((x/13.810381)^3.413225), geom="line", color = "#00BFC4") +
  stat_function(fun=function(x)exp(1)^-((x/11.529975)^6.687168), geom="line", color = "#00BFC4") +
  stat_function(fun=function(x)exp(1)^-((x/12.254437)^5.311160), geom="line", color = "#00BFC4") +
  stat_function(fun=function(x)exp(1)^-((x/13.515178)^4.590158), geom="line", color = "#00BFC4") +
  stat_function(fun=function(x)exp(1)^-((x/8.860129)^1.748245), geom="line", color = "#F8766D") +
  stat_function(fun=function(x)exp(1)^-((x/12.912245)^2.061356), geom="line", color = "#F8766D") +
  stat_function(fun=function(x)exp(1)^-((x/18.379138)^3.214658), geom="line", color = "#F8766D") +
  stat_function(fun=function(x)exp(1)^-((x/16.754905)^3.991739), geom="line", color = "#F8766D") +
  stat_function(fun=function(x)exp(1)^-((x/7.499330)^1.444372), geom="line", color = "#F8766D") +
  stat_function(fun=function(x)exp(1)^-((x/10.286655)^1.566060), geom="line", color = "#F8766D") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 12, color = "black"),
        legend.text = element_text(size = 12, color = "black")) + 
  labs(title = "", color = "", x = "Day", y = "Fraction Alive") +
  theme(legend.position="right")
n2_surv_fit
# find colors for 2 groups. Enter 2 below, 8 groups enter 8
# show_col(hue_pal()(8))

# Effect of shape and scale parameters, N2 has large shape parameter b/c this is capturing variation in mean lifespan. 
 weibull_parameter_space <- ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  xlim(0,30) +
  ylim(0,1) +
  stat_function(fun=function(x)exp(1)^-((x/5)^1), geom="line", color = "red") +
  stat_function(fun=function(x)exp(1)^-((x/10)^1), geom="line", color = "orange") +
  stat_function(fun=function(x)exp(1)^-((x/20)^1), geom="line", color = "green") +
  stat_function(fun=function(x)exp(1)^-((x/1)^1), geom="line", color = "blue") +
  stat_function(fun=function(x)exp(1)^-((x/10)^2.5), geom="line", color = "purple") +
  stat_function(fun=function(x)exp(1)^-((x/10)^5), geom="line", color = "pink") +
  stat_function(fun=function(x)exp(1)^-((x/5)^2.5), geom="line", color = "purple") +
  stat_function(fun=function(x)exp(1)^-((x/5)^5), geom="line", color = "pink") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 12, color = "black"),
        legend.text = element_text(size = 12, color = "black")) + 
  labs(title = "Weibull parameter space", color = "plate", x = "Day", y = "Fraction Alive") +
  theme(legend.position="right")
#weibull_parameter_space
 
```
<br> Large N2 shape parameter is largely driven by rapid decline in survival relative to other strains.

###Calculate H2 for ALL blocks (1-4) and all traits with block regression
```{r, echo=FALSE, warnings=FALSE}
# do block regression and calculate H2 for all traits.
old_dat <- as.data.frame(read.csv(file = "/Users/tim/Documents/WorMotel/2017_Pilot1_Survival/20170116_FullProcessed.csv", header = T)) %>%
  dplyr::rename(plate_num = bio_rep, replicate_num = tech_rep, worm_num = rep, lifespan_days = life_span, time_days = time_d) %>%
  dplyr::mutate(dataset = "old")
  
new_dat <- as.data.frame(read.csv(file = "WildIsolateLSData_180515.csv", header = T)) %>%
  dplyr::mutate(lifespan_days = ifelse(lifespan_days == "NaN", NA, lifespan_days),
                plate_num = plate_num + 6,
                replicate_num = replicate_num + 2,
                activity_type = as.factor(ifelse(activity_type == -1, "stim", "base")),
                dataset = "new",
                activity = ifelse(activity == "NaN", NA, activity))

full_dat <- dplyr::bind_rows(old_dat, new_dat)

# ADD mean survival, fit weibull, add activity traits to full dataset.
full_dat_proc <- full_dat %>%
 dplyr::filter(!is.na(lifespan_days)) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(mean_ls = mean(lifespan_days),
                cv_ls = sd(lifespan_days, na.rm = T)/ 
   mean(lifespan_days, na.rm = T)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(plate_num = factor(plate_num),
                replicate_num = factor(replicate_num),
                worm_num = factor(worm_num),
                dataset = factor(dataset))
  

# reshape data for survival object
full_dat_w <- full_dat %>%
  dplyr::distinct(strain, plate_num, worm_num, lifespan_days) %>%
  dplyr::select(strain, plate_num, worm_num, lifespan_days) %>%
  dplyr::mutate(status = ifelse(is.na(lifespan_days), 0, 1)) %>%
  dplyr::mutate(plate_num = factor(plate_num))

#Asign functions to calculate 90, 50, 10 percent survival times with Weibull model.
weibull_50 <- function(shape, scale) {
  qweibull(0.5, shape = shape, scale = scale)
}
weibull_10 <- function(shape, scale) {
  qweibull(0.9, shape = shape, scale = scale)
}
weibull_90 <- function(shape, scale) {
  qweibull(0.1, shape = shape, scale = scale)
}

# establish lists to populate with for loop
block_list <- list()
strain_list <- list()

# Calculate weibull shape and scale parameters for all strains in each plate
for(b in 1:length(unique(full_dat_w$plate_num))){
  
  block_df <- full_dat_w %>%
    dplyr::filter(plate_num == (unique(full_dat_w$plate_num)[b]))

  for(s in 1:length(unique(block_df$strain))){
  
    strain_df <- block_df %>%
      dplyr::filter(strain == (unique(block_df$strain)[s]))
    
    if(sum(strain_df$status) != 0){
    #run weibull fit and save parameters
    w_fit = flexsurvreg(Surv(lifespan_days, status) ~ 1, dist = "weibull", data = strain_df)
  
    w_90 <- summary(w_fit, fn = weibull_90, t=1)
  
    w_50 <- summary(w_fit, fn = weibull_50, t=1)
  
    w_10 <- summary(w_fit, fn = weibull_10, t=1)
  
    strain_surv <- data.frame(plate_num = unique(block_df$plate_num),
                  strain = unique(block_df$strain)[s],
                  shape = w_fit[["res"]][1],
                  scale = w_fit[["res"]][2],
                  LT10 = w_90[[1]][2],
                  LT50 = w_50[[1]][2],
                  LT90 = w_10[[1]][2])
    strain_surv <- strain_surv %>%
      dplyr::rename('LT10' = est, 'LT50' = est.1, 'LT90' = est.2)

    strain_list[[s]] <- strain_surv 

    } else {
      #fill in strain surv with NAs
      strain_surv <- data.frame(plate_num = unique(block_df$plate_num),
                  strain = unique(block_df$strain)[s],
                  shape = NA,
                  scale = NA,
                  LT10 = NA,
                  LT50 = NA,
                  LT90 = NA)

    strain_list[[s]] <- strain_surv 
  }
  }
  
  strain_df <- bind_rows(strain_list)
  
  block_list[[b]] <- strain_df
}

#combine all data
full_h_df <- dplyr::bind_rows(block_list) 

full_dat_proc_2 <- left_join(full_dat_proc, full_h_df)

# All blocks with block regression and calculate H2. dropping second day 22.5 value from plates 11 and 12. Need to check with Matt if this is right thing to do. It appears 22.5 was measured twice. second value is usually 0 so maybe this one should be dropped.
full_dat_proc_3 <- full_dat_proc_2 %>%
  dplyr::group_by(strain, replicate_num, plate_num, activity_type, worm_num) %>%
  dplyr::distinct(time_days, .keep_all = T) %>% # remove extra 22.5 days timepoint, may need to do this differently. Confirm with Matt Churgin
  dplyr::ungroup() %>%
  tidyr::spread(key = activity_type, value = activity) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(base_outliers = remove_outliers(base),
                stim_outliers = remove_outliers(stim)) %>% # this is pretty aggressive trimming, can cut back to under 1000 maybe?
  dplyr::group_by(replicate_num, plate_num, strain, worm_num) %>%
  dplyr::mutate(w_int_b_activity =  sum(base_outliers, na.rm = T),
                w_int_s_activity = sum(stim_outliers, na.rm = T),
                w_norm_int_b_activity = (w_int_b_activity/lifespan_days),
                w_norm_int_s_activity = (w_int_s_activity/lifespan_days)) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(mean_int_b_activity =  mean(w_int_b_activity, na.rm = T),
                mean_int_s_activity = mean(w_int_s_activity, na.rm = T),
                mean_norm_int_b_activity = (mean_int_b_activity/lifespan_days),
                mean_norm_int_s_activity = (mean_int_b_activity/lifespan_days)) %>%
  dplyr::ungroup() %>%
  dplyr::select(-time_days, -base, -stim, -base_outliers, -stim_outliers, -dataset) %>%# removing unused time data
  dplyr::distinct(strain, replicate_num, plate_num, worm_num, .keep_all = T) %>%
  tidyr::gather(trait, phenotype, -strain, -replicate_num, -plate_num, -worm_num)

# individual worm traits (5)
ind_w_traits <- full_dat_proc_3 %>%
  dplyr::filter(trait %in% c("lifespan_days", "w_int_b_activity", "w_int_s_activity", "w_norm_int_b_activity", "w_norm_int_s_activity")) %>%
  dplyr::group_by(trait) %>%
  dplyr::mutate(residual_phenotype = residuals(lm(phenotype ~ replicate_num))) %>%
  dplyr::ungroup()

# mean worMotel traits (11)
mean_traits <- full_dat_proc_3 %>%
  dplyr::filter(!trait %in% c("lifespan_days", "w_int_b_activity", "w_int_s_activity", "w_norm_int_b_activity", "w_norm_int_s_activity")) %>%
  dplyr::distinct(strain, replicate_num, plate_num, trait, .keep_all = T) %>%
  dplyr::select(-worm_num) %>%
  dplyr::group_by(trait) %>%
  dplyr::mutate(residual_phenotype = residuals(lm(phenotype ~ replicate_num))) %>%
  dplyr::ungroup()


# plot raw data for each individual trait without block regression
ind_w_traits_plot <- ggplot(ind_w_traits) +
  geom_jitter(aes(x=strain, y = phenotype, color = replicate_num), width = 0.25) +
  geom_boxplot(aes(x=strain, y = phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Individual traits", color = "block", x = "", y = "Phenotype")
ind_w_traits_plot

# plot raw data for each mean trait without block regression
mean_traits_plot <- ggplot(mean_traits) +
  geom_jitter(aes(x=strain, y = phenotype, color = replicate_num), width = 0.25) +
  geom_boxplot(aes(x=strain, y = phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Mean traits", color = "block", x = "", y = "Phenotype")
mean_traits_plot

# individual traits regress out block effect
ind_w_traits_r_plot <- ggplot(ind_w_traits) +
  geom_jitter(aes(x=strain, y = residual_phenotype, color = replicate_num), width = 0.25) +
  geom_boxplot(aes(x=strain, y = residual_phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Individual traits block regressed", color = "block", x = "", y = "residual_phenotype")
ind_w_traits_r_plot

# mean traitis regress out block effect
mean_traits_r_plot <- ggplot(mean_traits) +
  geom_jitter(aes(x=strain, y = residual_phenotype, color = replicate_num), width = 0.25) +
  geom_boxplot(aes(x=strain, y = residual_phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Mean traits block regressed", color = "block", x = "", y = "residual_Phenotype")
mean_traits_r_plot

# calculate heritability for raw and block regressed traits

```

###Calculate H2 within each of new blocks
```{r, echo=FALSE, warnings=FALSE}


```

###H2 for activity
```{r, echo=FALSE, warnings=FALSE}


```

###H2 for mean lifespan variation (CV)
```{r, echo=FALSE, warnings=FALSE}


```
