---
title: "WorMotel lifespan Heritability"
author: "Tim C."
date: "5/29/2018"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---
This code uses processed survival data from Chris Fang-Yen's lab to calculate heritability of survival traits. The data were collected with WorMotel in four replicate blocks by Matt Churgin at UPenn. Four independent bleaches of 12 strains (blocks) were run on 2-4 replicate wormmotels (20 nematodes / strain). 

```{r, echo=F, warning = F, message=F}
library(rio)
library(ggplot2)
library(dplyr)
library(survival)
library(flexsurv)
library(broom)
library(minpack.lm)
library(forcats)
library(glue)
library(scales)
library(boot)
try(setwd(dirname(rstudioapi::getActiveDocumentContext()$path)))
```

```{r, echo=FALSE, warning=FALSE}
# define heritability test
# data is data frame that contains strain and phenotype column
# indicies are used by the boot function to sample from the 'data' data.frame
H2.test.boot <- function(data, indicies){

  d <- data[indicies,]

  pheno <- as.data.frame(dplyr::select(d, phenotype))[,1]
  strain <- as.factor(d$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}
# data is data frame that contains strain and phenotype column
H2.test <- function(data){

  pheno <- as.data.frame(dplyr::select(data, phenotype))[,1]
  strain <- as.factor(data$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}

# df is data frame that contains strain and phenotype column
H2.calc <- function(df, boot = T){


  if(boot == T){
    # bootstrapping with 1000 replications
    results <- boot(data=df, statistic=H2.test.boot, R=10000)

    # get 95% confidence interval
    ci <- boot.ci(results, type="bca")

    H2_errors <- data.frame(H2 = ci$t0, ci_l = ci$bca[4], ci_r = ci$bca[5])

    return(H2_errors)

  } else {

    H2 <- data.frame(H2 = H2.test(data = df), ci_l = NA, ci_r = NA)
    return(H2)
  }

}


# remove outliers function
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}
```

## Trait heritability (all blocks)
```{r, echo=FALSE, warnings=FALSE}
# do block regression and calculate H2 for all traits.
old_dat <- as.data.frame(read.csv(file = "/Users/tim/Documents/WorMotel/2017_Pilot1_Survival/20170116_FullProcessed.csv", header = T)) %>%
  dplyr::rename(plate_num = bio_rep, replicate_num = tech_rep, worm_num = rep, lifespan_days = life_span, time_days = time_d) %>%
  dplyr::mutate(dataset = "old")
  
new_dat <- as.data.frame(read.csv(file = "WildIsolateLSData_180515.csv", header = T)) %>%
  dplyr::mutate(lifespan_days = ifelse(lifespan_days == "NaN", NA, lifespan_days),
                plate_num = plate_num + 6,
                replicate_num = replicate_num + 2,
                activity_type = as.factor(ifelse(activity_type == -1, "stim", "base")),
                dataset = "new",
                activity = ifelse(activity == "NaN", NA, activity))

full_dat <- dplyr::bind_rows(old_dat, new_dat)

# ADD mean survival, fit weibull, add activity traits to full dataset.
full_dat_proc <- full_dat %>%
 dplyr::filter(!is.na(lifespan_days)) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(mean_ls = mean(lifespan_days),
                cv_ls = sd(lifespan_days, na.rm = T)/ 
   mean(lifespan_days, na.rm = T)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(plate_num = factor(plate_num),
                replicate_num = factor(replicate_num),
                worm_num = factor(worm_num),
                dataset = factor(dataset))
  

# reshape data for survival object
full_dat_w <- full_dat %>%
  dplyr::distinct(strain, plate_num, worm_num, lifespan_days) %>%
  dplyr::select(strain, plate_num, worm_num, lifespan_days) %>%
  dplyr::mutate(status = ifelse(is.na(lifespan_days), 0, 1)) %>%
  dplyr::mutate(plate_num = factor(plate_num))

#Asign functions to calculate 90, 50, 10 percent survival times with Weibull model.
weibull_50 <- function(shape, scale) {
  qweibull(0.5, shape = shape, scale = scale)
}
weibull_10 <- function(shape, scale) {
  qweibull(0.9, shape = shape, scale = scale)
}
weibull_90 <- function(shape, scale) {
  qweibull(0.1, shape = shape, scale = scale)
}

# establish lists to populate with for loop
block_list <- list()
strain_list <- list()

# Calculate weibull shape and scale parameters for all strains in each plate
for(b in 1:length(unique(full_dat_w$plate_num))){
  
  block_df <- full_dat_w %>%
    dplyr::filter(plate_num == (unique(full_dat_w$plate_num)[b]))

  for(s in 1:length(unique(block_df$strain))){
  
    strain_df <- block_df %>%
      dplyr::filter(strain == (unique(block_df$strain)[s]))
    
    if(sum(strain_df$status) != 0){
    #run weibull fit and save parameters
    w_fit = flexsurvreg(Surv(lifespan_days, status) ~ 1, dist = "weibull", data = strain_df)
  
    w_90 <- summary(w_fit, fn = weibull_90, t=1)
  
    w_50 <- summary(w_fit, fn = weibull_50, t=1)
  
    w_10 <- summary(w_fit, fn = weibull_10, t=1)
  
    strain_surv <- data.frame(plate_num = unique(block_df$plate_num),
                  strain = unique(block_df$strain)[s],
                  shape = w_fit[["res"]][1],
                  scale = w_fit[["res"]][2],
                  LT10 = w_90[[1]][2],
                  LT50 = w_50[[1]][2],
                  LT90 = w_10[[1]][2])
    strain_surv <- strain_surv %>%
      dplyr::rename('LT10' = est, 'LT50' = est.1, 'LT90' = est.2)

    strain_list[[s]] <- strain_surv 

    } else {
      #fill in strain surv with NAs
      strain_surv <- data.frame(plate_num = unique(block_df$plate_num),
                  strain = unique(block_df$strain)[s],
                  shape = NA,
                  scale = NA,
                  LT10 = NA,
                  LT50 = NA,
                  LT90 = NA)

    strain_list[[s]] <- strain_surv 
  }
  }
  
  strain_df <- bind_rows(strain_list)
  
  block_list[[b]] <- strain_df
}

#combine all data
full_h_df <- dplyr::bind_rows(block_list) 

full_dat_proc_2 <- left_join(full_dat_proc, full_h_df)

# All blocks with block regression and calculate H2. dropping second day 22.5 value from plates 11 and 12. Need to check with Matt if this is right thing to do. It appears 22.5 was measured twice. second value is usually 0 so maybe this one should be dropped.
full_dat_proc_3 <- full_dat_proc_2 %>%
  dplyr::group_by(strain, replicate_num, plate_num, activity_type, worm_num) %>%
  dplyr::distinct(time_days, .keep_all = T) %>% # remove extra 22.5 days timepoint, may need to do this differently. Confirm with Matt Churgin
  dplyr::ungroup() %>%
  tidyr::spread(key = activity_type, value = activity) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(base_outliers = remove_outliers(base),
                stim_outliers = remove_outliers(stim)) %>% # this is pretty aggressive trimming, can cut back to under 1000 maybe?
  dplyr::group_by(replicate_num, plate_num, strain, worm_num) %>%
  dplyr::mutate(w_int_b_activity =  sum(base_outliers, na.rm = T),
                w_int_s_activity = sum(stim_outliers, na.rm = T),
                w_norm_int_b_activity = (w_int_b_activity/lifespan_days),
                w_norm_int_s_activity = (w_int_s_activity/lifespan_days)) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(mean_int_b_activity =  mean(w_int_b_activity, na.rm = T),
                mean_int_s_activity = mean(w_int_s_activity, na.rm = T),
                mean_norm_int_b_activity = (mean_int_b_activity/lifespan_days),
                mean_norm_int_s_activity = (mean_int_b_activity/lifespan_days)) %>%
  dplyr::ungroup() %>%
  dplyr::select(-time_days, -base, -stim, -base_outliers, -stim_outliers, -dataset) %>%# removing unused time data
  dplyr::distinct(strain, replicate_num, plate_num, worm_num, .keep_all = T) %>%
  tidyr::gather(trait, phenotype, -strain, -replicate_num, -plate_num, -worm_num)

# individual worm traits (5)
ind_w_traits <- full_dat_proc_3 %>%
  dplyr::filter(trait %in% c("lifespan_days", "w_int_b_activity", "w_int_s_activity", "w_norm_int_b_activity", "w_norm_int_s_activity")) %>%
  dplyr::group_by(trait) %>%
  dplyr::mutate(residual_phenotype = residuals(lm(phenotype ~ replicate_num))) %>%
  dplyr::ungroup()

# mean worMotel traits (11)
mean_traits <- full_dat_proc_3 %>%
  dplyr::filter(!trait %in% c("lifespan_days", "w_int_b_activity", "w_int_s_activity", "w_norm_int_b_activity", "w_norm_int_s_activity")) %>%
  dplyr::distinct(strain, replicate_num, plate_num, trait, .keep_all = T) %>%
  dplyr::select(-worm_num) %>%
  dplyr::group_by(trait) %>%
  dplyr::mutate(residual_phenotype = residuals(lm(phenotype ~ replicate_num))) %>%
  dplyr::ungroup()


# plot raw data for each individual trait without block regression
ind_w_traits_plot <- ggplot(ind_w_traits) +
  geom_jitter(aes(x=strain, y = phenotype, color = replicate_num), width = 0.25) +
  geom_boxplot(aes(x=strain, y = phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Individual traits", color = "block", x = "", y = "Phenotype")


# plot raw data for each mean trait without block regression
mean_traits_plot <- ggplot(mean_traits) +
  geom_jitter(aes(x=strain, y = phenotype, color = replicate_num), width = 0.25, size = 0.5) +
  geom_boxplot(aes(x=strain, y = phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Fig 1: Traits", color = "block", x = "", y = "Value")
mean_traits_plot

# individual traits regress out block effect
ind_w_traits_r_plot <- ggplot(ind_w_traits) +
  geom_jitter(aes(x=strain, y = residual_phenotype, color = replicate_num), width = 0.25) +
  geom_boxplot(aes(x=strain, y = residual_phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Individual traits block regressed", color = "block", x = "", y = "residual_phenotype")


# mean traitis regress out block effect
mean_traits_r_plot <- ggplot(mean_traits) +
  geom_jitter(aes(x=strain, y = residual_phenotype, color = replicate_num), width = 0.25, size = 0.5) +
  geom_boxplot(aes(x=strain, y = residual_phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Fig 2: Traits block regressed", color = "block", x = "", y = "residual_value")
mean_traits_r_plot

# calculate heritability for raw and block regressed traits
H2_ind_traits <- ind_w_traits %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

H2_ind_traits_r <- ind_w_traits %>%
  dplyr::mutate(phenotype = residual_phenotype) %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

H2_mean_traits <- mean_traits %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)%>%
  dplyr::ungroup()

H2_mean_traits_r <- mean_traits %>%
  dplyr::mutate(phenotype = residual_phenotype) %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

# calculate per block (replicate number) then do heritability
H2_mean_traits_block <- mean_traits %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(strain, replicate_num, trait) %>%
  dplyr::mutate(phenotype = mean(phenotype)) %>%
  dplyr::select(-plate_num) %>%
  dplyr::distinct(strain, replicate_num, trait, .keep_all = T) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

H2_mean_traits_r_block <- mean_traits %>%
  dplyr::mutate(phenotype = residual_phenotype) %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(strain, replicate_num, trait) %>%
  dplyr::mutate(phenotype = mean(phenotype)) %>%
  dplyr::select(-plate_num) %>%
  dplyr::distinct(strain, replicate_num, trait, .keep_all = T) %>%
  dplyr::group_by(trait)
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()


# plot heritability for all traits
h2_ind_traits_plot <- ggplot() +
  geom_point(data = H2_ind_traits, aes(x = trait, y = H2, color = "raw")) +
  geom_point(data = H2_ind_traits_r, aes(x = trait, y = H2, color = "block regressed")) +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Individual traits", color = "", x = "", y = "Heritability")

h2_mean_traits_plot <- ggplot() +
  geom_jitter(data = H2_mean_traits, aes(x = trait, y = H2, color = "raw"), width = .1) +
  geom_jitter(data = H2_mean_traits_r, aes(x = trait, y = H2, color = "block regressed"), width = .1) +
  #geom_jitter(data = H2_mean_traits_r_block, aes(x = trait, y = H2, color = "plate means block regressed"), width = .1) +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Fig 3: Heritability of traits", color = "", x = "", y = "Heritability")
h2_mean_traits_plot
```
<br>
`cv_ls` = coefficient of variation for lifespan <br>
`LT10` = time to 10% death <br>
`LT50` = time to 50% death <br>
`LT90` = time to 90% death <br>
`mean_int_b_activity` = mean integrated base activity <br>
`mean_int_s_activity` = mean integrated stimulated activity <br>
`mean_ls = mean lifespan` <br>
`mean_norm_int_b_activity` = mean integrated base activity / mean lifespan <br>
`mean_norm_int_s_activity` = mean integrated stimulated activity / mean lifespan <br>
`scale` = survival fit parameter <br>
`shape` = survival fit parameter <br>

## Trait heritability (blocks 3 and 4 only)
```{r, echo=FALSE, warnings=FALSE}
# filter to new dataset
full_dat_proc_new <- full_dat_proc_2 %>%
  dplyr::filter(dataset == "new") %>%
  dplyr::group_by(strain, replicate_num, plate_num, activity_type, worm_num) %>%
  dplyr::distinct(time_days, .keep_all = T) %>% # remove extra 22.5 days timepoint, may need to do this differently. Confirm with Matt Churgin
  dplyr::ungroup() %>%
  tidyr::spread(key = activity_type, value = activity) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(base_outliers = remove_outliers(base),
                stim_outliers = remove_outliers(stim)) %>% # this is pretty aggressive trimming, can cut back to under 1000 maybe?
  dplyr::group_by(replicate_num, plate_num, strain, worm_num) %>%
  dplyr::mutate(w_int_b_activity =  sum(base_outliers, na.rm = T),
                w_int_s_activity = sum(stim_outliers, na.rm = T),
                w_norm_int_b_activity = (w_int_b_activity/lifespan_days),
                w_norm_int_s_activity = (w_int_s_activity/lifespan_days)) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(mean_int_b_activity =  mean(w_int_b_activity, na.rm = T),
                mean_int_s_activity = mean(w_int_s_activity, na.rm = T),
                mean_norm_int_b_activity = (mean_int_b_activity/lifespan_days),
                mean_norm_int_s_activity = (mean_int_b_activity/lifespan_days)) %>%
  dplyr::ungroup() %>%
  dplyr::select(-time_days, -base, -stim, -base_outliers, -stim_outliers, -dataset) %>%# removing unused time data
  dplyr::distinct(strain, replicate_num, plate_num, worm_num, .keep_all = T) %>%
  tidyr::gather(trait, phenotype, -strain, -replicate_num, -plate_num, -worm_num)

# mean worMotel traits (11)
mean_traits_new <- full_dat_proc_new %>%
  dplyr::filter(!trait %in% c("lifespan_days", "w_int_b_activity", "w_int_s_activity", "w_norm_int_b_activity", "w_norm_int_s_activity")) %>%
  dplyr::distinct(strain, replicate_num, plate_num, trait, .keep_all = T) %>%
  dplyr::select(-worm_num) %>%
  dplyr::group_by(trait) %>%
  dplyr::mutate(residual_phenotype = residuals(lm(phenotype ~ replicate_num))) %>%
  dplyr::ungroup()

# plot raw data for each mean trait without block regression
mean_traits_new_plot <- ggplot(mean_traits_new) +
  geom_jitter(aes(x=strain, y = phenotype, color = replicate_num), width = 0.25, size = 0.5) +
  geom_boxplot(aes(x=strain, y = phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Fig 4: Traits", color = "block", x = "", y = "Value")
mean_traits_new_plot

# mean traitis regress out block effect
mean_traits_new_r_plot <- ggplot(mean_traits_new) +
  geom_jitter(aes(x=strain, y = residual_phenotype, color = replicate_num), width = 0.25, size = 0.5) +
  geom_boxplot(aes(x=strain, y = residual_phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Fig 5: Traits block regressed", color = "block", x = "", y = "residual_value")
mean_traits_new_r_plot

#heritability for new
H2_mean_traits_new <- mean_traits_new %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)%>%
  dplyr::ungroup()

H2_mean_traits_new_r <- mean_traits_new %>%
  dplyr::mutate(phenotype = residual_phenotype) %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

# calculate per block (replicate number) then do heritability
H2_mean_traits_new_block <- mean_traits_new %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(strain, replicate_num, trait) %>%
  dplyr::mutate(phenotype = mean(phenotype)) %>%
  dplyr::select(-plate_num) %>%
  dplyr::distinct(strain, replicate_num, trait, .keep_all = T) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

H2_mean_traits_r_new_block <- mean_traits_new %>%
  dplyr::mutate(phenotype = residual_phenotype) %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::group_by(strain, replicate_num, trait) %>%
  dplyr::mutate(phenotype = mean(phenotype)) %>%
  dplyr::select(-plate_num) %>%
  dplyr::distinct(strain, replicate_num, trait, .keep_all = T) %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

#drop N2 and test again
#heritability for new
H2_mean_traits_new_no_n2 <- mean_traits_new %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::filter(strain != "N2") %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)%>%
  dplyr::ungroup()

H2_mean_traits_new_r_no_n2 <- mean_traits_new %>%
  dplyr::mutate(phenotype = residual_phenotype) %>%
  dplyr::select(-residual_phenotype) %>%
  dplyr::filter(strain != "N2") %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

# plot heritability 
h2_mean_traits_new_plot <- ggplot() +
  geom_jitter(data = H2_mean_traits_new, aes(x = trait, y = H2, color = "raw"), width = .1) +
  geom_jitter(data = H2_mean_traits_new_r, aes(x = trait, y = H2, color = "block regressed"), width = .1) +
  #geom_jitter(data = H2_mean_traits_r_new_block, aes(x = trait, y = H2, color = "plate means block regressed"), width = .1) +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Fig 6: Heritability of traits (all strains)", color = "", x = "", y = "Heritability")
h2_mean_traits_new_plot

# plot heritability no N2 
h2_mean_traits_new_no_n2_plot <- ggplot() +
  geom_jitter(data = H2_mean_traits_new_no_n2, aes(x = trait, y = H2, color = "raw"), width = .1) +
  geom_jitter(data = H2_mean_traits_new_r_no_n2, aes(x = trait, y = H2, color = "block regressed"), width = .1) +
  #geom_jitter(data = H2_mean_traits_r_new_block, aes(x = trait, y = H2, color = "plate means block regressed"), width = .1) +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Fig 7: Heritability of traits (N2  excluded)", color = "", x = "", y = "Heritability")
h2_mean_traits_new_no_n2_plot
```
<br>
1) Restricting analysis to new data only results in higher H2 values (Fig. 3 vs. Fig. 6). <br>
2) N2 is driving the high heritability of `cv_ls` and `shape` traits (Fig. 6 vs. Fig. 7). <br>
3) JU775 and DL238 have high integrated activity traits (Fig. 5). This is consistent with observations in the past. <br>
4) How can we best normalize activity traits to capture healthspan? Normalization by mean_lifespan seems odd.

## Subsampling within blocks
Randomly subsample a single worMotel for each block 10 times and calculate H2 w/ block regression.<br><br>
```{r, echo=FALSE, warnings=FALSE}
# subsample single worMotel for each block 10 times and calculate H2
dat_subsample <- full_dat_proc_2 %>%
  dplyr::group_by(strain, replicate_num, plate_num, activity_type, worm_num) %>%
  dplyr::distinct(time_days, .keep_all = T) %>% # remove extra 22.5 days timepoint, may need to do this differently. Confirm with Matt Churgin
  dplyr::ungroup() %>%
  tidyr::spread(key = activity_type, value = activity) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(base_outliers = remove_outliers(base),
                stim_outliers = remove_outliers(stim)) %>%# this is pretty aggressive trimming, can cut back to under 1000 maybe?
  dplyr::group_by(replicate_num, plate_num, strain, worm_num) %>%
  dplyr::mutate(w_int_b_activity =  sum(base_outliers, na.rm = T),
                w_int_s_activity = sum(stim_outliers, na.rm = T),
                w_norm_int_b_activity = (w_int_b_activity/lifespan_days),
                w_norm_int_s_activity = (w_int_s_activity/lifespan_days)) %>%
  dplyr::ungroup() %>%
  dplyr::select(-time_days, -base, -stim, -base_outliers, -stim_outliers, -dataset, -lifespan_days) %>% # removing unused time data
  dplyr::distinct(strain, replicate_num, plate_num, worm_num, .keep_all = T) %>%
  dplyr::group_by(replicate_num, plate_num, strain) %>%
  dplyr::mutate(mean_int_b_activity =  mean(w_int_b_activity, na.rm = T),
                mean_int_s_activity = mean(w_int_s_activity, na.rm = T),
                mean_norm_int_b_activity = mean(w_norm_int_b_activity),
                mean_norm_int_s_activity = mean(w_norm_int_s_activity)) %>%
  dplyr::select(-w_int_b_activity, -w_int_s_activity, -w_norm_int_b_activity, -w_norm_int_s_activity, -worm_num) %>%
  dplyr::distinct(strain, replicate_num, plate_num, .keep_all = T) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(plate_num = case_when(
                                      plate_num == "3" ~ 1,
                                      plate_num == "5" ~ 2,
                                      plate_num == "1" ~ 1,
                                      plate_num == "2" ~ 2,
                                      plate_num == "4" ~ 3,
                                      plate_num == "6" ~ 4,
                                      plate_num == "7" ~ 1,
                                      plate_num == "8" ~ 2,
                                      plate_num == "9" ~ 3,
                                      plate_num == "10" ~ 1,
                                      plate_num == "11" ~ 2,
                                      plate_num == "12" ~ 3))


# attempt at subsampling
random_plates <- as.data.frame(cbind(as.vector(sample(1:2, 100, replace=T)),
            as.vector(sample(1:4, 100, replace=T)),
             as.vector(sample(1:3, 100, replace=T)),
            as.vector(sample(1:3, 100, replace=T)))) %>% # only allowing plate 1 for this rep b/c samples missing. Need to fix
  distinct() %>%
  .[1:10,]

ss1 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[1]) |
                (replicate_num == "2" & plate_num == random_plates$V2[1]) |
                (replicate_num == "3" & plate_num == random_plates$V3[1]) |
                (replicate_num == "4" & plate_num == random_plates$V4[1])) %>%
  dplyr::mutate(sample = 1)
ss2 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[2]) |
                (replicate_num == "2" & plate_num == random_plates$V2[2]) |
                (replicate_num == "3" & plate_num == random_plates$V3[2]) |
                (replicate_num == "4" & plate_num == random_plates$V4[2])) %>%
  dplyr::mutate(sample = 2)
ss3 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[3]) |
                (replicate_num == "2" & plate_num == random_plates$V2[3]) |
                (replicate_num == "3" & plate_num == random_plates$V3[3]) |
                (replicate_num == "4" & plate_num == random_plates$V4[3])) %>%
  dplyr::mutate(sample = 3)
ss4 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[4]) |
                (replicate_num == "2" & plate_num == random_plates$V2[4]) |
                (replicate_num == "3" & plate_num == random_plates$V3[4]) |
                (replicate_num == "4" & plate_num == random_plates$V4[4])) %>%
  dplyr::mutate(sample = 4)
ss5 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[5]) |
                (replicate_num == "2" & plate_num == random_plates$V2[5]) |
                (replicate_num == "3" & plate_num == random_plates$V3[5]) |
                (replicate_num == "4" & plate_num == random_plates$V4[5])) %>%
  dplyr::mutate(sample = 5)
ss6 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[6]) |
                (replicate_num == "2" & plate_num == random_plates$V2[6]) |
                (replicate_num == "3" & plate_num == random_plates$V3[6]) |
                (replicate_num == "4" & plate_num == random_plates$V4[6])) %>%
  dplyr::mutate(sample = 6)
ss7 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[7]) |
                (replicate_num == "2" & plate_num == random_plates$V2[7]) |
                (replicate_num == "3" & plate_num == random_plates$V3[7]) |
                (replicate_num == "4" & plate_num == random_plates$V4[7])) %>%
  dplyr::mutate(sample = 7)
ss8 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[8]) |
                (replicate_num == "2" & plate_num == random_plates$V2[8]) |
                (replicate_num == "3" & plate_num == random_plates$V3[8]) |
                (replicate_num == "4" & plate_num == random_plates$V4[8])) %>%
  dplyr::mutate(sample = 8)
ss9 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[9]) |
                (replicate_num == "2" & plate_num == random_plates$V2[9]) |
                (replicate_num == "3" & plate_num == random_plates$V3[9]) |
                (replicate_num == "4" & plate_num == random_plates$V4[9])) %>%
  dplyr::mutate(sample = 9)
ss10 <- dat_subsample %>%
  dplyr::filter((replicate_num == "1" & plate_num == random_plates$V1[10]) |
                (replicate_num == "2" & plate_num == random_plates$V2[10]) |
                (replicate_num == "3" & plate_num == random_plates$V3[10]) |
                (replicate_num == "4" & plate_num == random_plates$V4[10])) %>%
  dplyr::mutate(sample = 10)

# calculate heritability on subsampled data
ss_all_gather <- bind_rows(ss1, ss2, ss3, ss4, ss5, ss6, ss7, ss8, ss9, ss10) %>%
  dplyr::group_by(sample) %>%
  dplyr::mutate(mean_ls = residuals(lm(mean_ls ~ replicate_num)),
                cv_ls = residuals(lm(cv_ls ~ replicate_num)),
                shape = residuals(lm(shape ~ replicate_num)),
                scale = residuals(lm(scale ~ replicate_num)),
                LT10 = residuals(lm(LT10 ~ replicate_num)),
                LT50 = residuals(lm(LT50 ~ replicate_num)),
                LT90 = residuals(lm(LT90 ~ replicate_num)),
                mean_int_b_activity = residuals(lm(mean_int_b_activity ~ replicate_num)),
                mean_int_s_activity = residuals(lm(mean_int_s_activity ~ replicate_num)),
                mean_norm_int_b_activity = residuals(lm(mean_norm_int_b_activity ~ replicate_num)),
                mean_norm_int_s_activity = residuals(lm(mean_norm_int_s_activity ~ replicate_num))) %>%
  dplyr::group_by(strain, sample) %>%
  tidyr::gather(trait, phenotype, -strain, -replicate_num, -plate_num, -sample) 
  
ss_all_h2 <- ss_all_gather %>%
  dplyr::group_by(sample, trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x) %>%
  dplyr::ungroup()

# plot phenotypes
subsample_plot_1 <- ggplot(ss_all_gather %>% dplyr::filter(sample == 1)) +
  geom_jitter(aes(x=strain, y = phenotype, color = as.factor(replicate_num)), width = 0.25, size = 0.5) +
  geom_boxplot(aes(x=strain, y = phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Randomly sample single worMotel from each block (1)", color = "block", x = "", y = "value")




#plot heritability on subsampled data
h2_subsample_plot <- ggplot(ss_all_h2) +
  geom_jitter(aes(x=trait, y = H2, color = as.factor(sample)), width = 0.25) +
  geom_boxplot(aes(x=trait, y = H2), outlier.shape = NA, alpha = 0.25) +
  theme_light() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 12, color = "black"),
        legend.text = element_text(size = 12, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Fig. 8: Subsample single worMotel from each block (10 times)", color = "subsample", x = "", y = "Heritability")
h2_subsample_plot

```
<br>
1) Are we better off with a single worMotel per block? <br>
2) Other subsampling strategies? Inference for each?


## 24-well Assay
These data are from 24-well plates not worMotels. I believe 20 L4 nematodes were added to each well; there are two replicate wells for each strain in a block.<br><br>
**lifespan** = This is an estimate based on 2% of total stimulated activity remaining in a particular well.<br>
**healthspan** = Similar to lifespan but for base activity?<br><br>
```{r, echo=FALSE, warning=FALSE}
# read in survival data from 24-well plates
dat_24 <- as.data.frame(read.csv(file = "/Users/tim/Documents/WorMotel/20180517_Assays1_6/AllDataLongFormat_WildIsolate24wellplateLifespan.csv", header = T)) %>%
  dplyr::rename(plate_num = Plate.Number,
                strain = Genotype,
                well = Well.Number,
                lifespan = Lifespan.Estimate..Days.,
                healthspan = Healthspan.Estimate..Days.,
                activity_type = Activity.Type,
                time_days = Time.Point..Days.,
                activity = Activity) %>%
  dplyr::mutate(activity_type = ifelse(activity_type == -1, "stim", "base")) %>%
  tidyr::spread(activity_type, activity)
  
dat_24_proc <- dat_24 %>%
  dplyr::distinct(plate_num, strain, well, .keep_all = T) %>%
  dplyr::select(-time_days, -base, -stim) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(residual_ls = residuals(lm(lifespan ~ plate_num)),
                residual_hs = residuals(lm(healthspan ~ plate_num))) %>%
  tidyr::gather(trait, phenotype, -plate_num, -strain, -well)

  
# plot traits
traits_24_facet <- ggplot(dat_24_proc %>% dplyr::filter(trait != "residual_hs", trait != "residual_ls")) +
  geom_jitter(aes(x=strain, y = phenotype, color = as.factor(plate_num)), width = 0.25, size = 2) +
  geom_boxplot(aes(x=strain, y = phenotype), outlier.shape = NA, alpha = 0.25) +
  facet_wrap(~trait, scales = "free") +
  theme_light() +
  theme(title = element_text(size = 16, color = "black", face = "bold"),
        axis.title = element_text(size = 16, color = "black", face = "bold"),
        axis.text = element_text(size = 16, color = "black"),
        legend.text = element_text(size = 16, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        strip.text.x = element_text(size = 16, colour = "black")) + 
  labs(title = "24_well traits", color = "block", x = "", y = "Value")
traits_24_facet

ggsave('figure/24well_phenotypes.pdf', width = 12, height = 8)
ggsave('figure/24well_phenotypes.png', width = 12, height = 8)

# calculate heritability for all traits
dat_24_proc_h2 <- dat_24_proc %>%
  dplyr::group_by(trait) %>%
  dplyr::do(broom::tidy(H2.test(.))) %>%
  dplyr::rename(H2 = x)%>%
  dplyr::ungroup()

# plot heritabilities
dat_24_proc_h2_plot <- ggplot(dat_24_proc_h2 %>% dplyr::filter(trait != "residual_hs", trait != "residual_ls")) +
  geom_point(aes(x = trait, y = H2), size = 5) +
  ylim(0,1) +
  theme_light() +
  theme(title = element_text(size = 16, color = "black", face = "bold"),
        axis.title = element_text(size = 16, color = "black", face = "bold"),
        axis.text = element_text(size = 16, color = "black"),
        legend.text = element_text(size = 16, color = "black"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  labs(title = "Heritability for 24_well traits", color = "", x = "", y = "Heritability")
dat_24_proc_h2_plot

ggsave('figure/24well_phenotypes.pdf', width = 12, height = 8)
ggsave('figure/24well_phenotypes.png', width = 12, height = 8)

```
<br>
1) Heritability is HIGH for lifespan and healthspan! <br>
2) Very small block effect, regressing out this effect does not improve heritability by much. <br>
3) How well do these values relate to actual lifespan? How much does that matter?