---
title: "WorMotel Life Span Heritability"
author: "Tim C."
date: "1/17/2018"
output: html_document
---

This code uses processed survival data from Chris Fang-Yen's lab to calcualte heritability of survival traits. The data were collected with WorMotel in winter 2017 by Matt Churgin at UPenn.  

```{r, echo=F, warning = F, message=F}
library(rio)
library(ggplot2)
library(dplyr)
library(survival)
library(flexsurv)
try(setwd(dirname(rstudioapi::getActiveDocumentContext()$path)))
```

#### Read in data
```{r}
df <- as.data.frame(read.csv(file = "20170116_FullProcessed.csv", header = T))
glimpse(df)
```

#### Plot raw data
looking for errors, block effects. Not certain how they are classifying technical rep or biological rep we should clarrify this.
```{r, echo=FALSE}
df_proc1 <- df %>%
  dplyr::filter(activity_type != "base", !is.na(life_span)) %>%
  dplyr::mutate(alive = ifelse(!is.na(activity), 1, 0)) %>%
  dplyr::group_by(strain, tech_rep, bio_rep, time_d) %>%
  dplyr::summarise(frac_alive = sum(alive)/n(),
                   n = n())

# plot survival curves for all strains
raw_plot_all <- ggplot(df_proc1) +
  aes(x=time_d, y=frac_alive, color=strain) +
  geom_line() +
  labs(x="Day", y="Fraction alive") +
  theme(legend.position = "right") +
  theme_grey() +
  facet_grid(bio_rep ~ tech_rep, labeller = label_both)

# plot survival curves for just N2 and CB4856
raw_plot_subset <- ggplot(df_proc1 %>% dplyr::filter(grepl("N2", strain) | grepl("CB4856", strain))) +
  aes(x=time_d, y=frac_alive, color=strain) +
  geom_line() +
  labs(x="Day", y="Fraction alive") +
  theme(legend.position = "right") +
  theme_grey() + 
  facet_grid(bio_rep ~ tech_rep, labeller = label_both)
raw_plot_all
raw_plot_subset
```

####fit weibull model to data and extract parameters to be used as traits
The code will correct for block effects. This is done by fitting a linear model to define the block effect for each trait then using the residuals from those fits as corrected traits.
```{r, echo=FALSE}

# reshape data for survival object, ignore technical replicate and include just biological reps
df_km <- df %>%
  dplyr::distinct(strain, bio_rep, rep, life_span) %>%
  dplyr::mutate(status = ifelse(is.na(life_span), 0, 1)) %>%
  dplyr::mutate(bio_rep = factor(bio_rep))

#Asign functions to calculate 90, 50, 10 percent survival times with Weibull model.
weibull_50 <- function(shape, scale) {
  qweibull(0.5, shape = shape, scale = scale)
}
weibull_10 <- function(shape, scale) {
  qweibull(0.9, shape = shape, scale = scale)
}
weibull_90 <- function(shape, scale) {
  qweibull(0.1, shape = shape, scale = scale)
}

# establish lists to populate with for loop
block_list <- list()
strain_list <- list()

# nested for loop to extract parameters for each strain in each block
for(b in 1:length(unique(df_km$bio_rep))){
  
  block_df <- df_km %>%
    dplyr::filter(bio_rep == (unique(df_km$bio_rep)[b]))

  for(s in 1:length(unique(block_df$strain))){
  
    strain_df <- block_df %>%
      dplyr::filter(strain == (unique(block_df$strain)[s]))
  
    w_fit = flexsurvreg(Surv(life_span, status) ~ 1, dist = "weibull", data = strain_df)
  
    w_90 <- summary(w_fit, fn = weibull_90, t=1)
  
    w_50 <- summary(w_fit, fn = weibull_50, t=1)
  
    w_10 <- summary(w_fit, fn = weibull_10, t=1)
  
    strain_surv <- data.frame(block = unique(block_df$bio_rep),
                  strain = unique(block_df$strain)[s],
                  w_shape_est = w_fit[["res"]][1],
                  w_scale_est = w_fit[["res"]][2],
                  w_surv90_est = w_90[[1]][2],
                  w_surv50_est = w_50[[1]][2],
                  w_surv10_est = w_10[[1]][2])
    strain_surv <- strain_surv %>%
      dplyr::rename('w_surv90_est' = est, 'w_surv50_est' = est.1, 'w_surv10_est' = est.2)

    strain_list[[s]] <- strain_surv
  }
  
  strain_df <- bind_rows(strain_list)
  
  block_list[[b]] <- strain_df
  
}

#combine all data
h_df <- dplyr::bind_rows(block_list) %>%
  dplyr::arrange(strain, block)

h_df_sh <- h_df %>%
  dplyr::rename(phenotype = w_shape_est)%>%
  dplyr::select(block, strain, phenotype)

shape_ris <- data.frame(strain = h_df_sh$strain, 
                        phenotype = h_df_sh$phenotype,
                        weibull_shape_residual =residuals(lm(phenotype ~ block, data = h_df_sh)))

h_df_sc <- h_df %>%
  dplyr::rename(phenotype = w_scale_est)%>%
  dplyr::select(block, strain, phenotype)

scale_ris <- data.frame(strain = h_df_sc$strain, 
                        phenotype = h_df_sc$phenotype,
                        weibull_scale_residual =residuals(lm(phenotype ~ block, data = h_df_sc)))

h_df_surv50 <- h_df %>%
  dplyr::rename(phenotype = w_surv50_est)%>%
  dplyr::select(block, strain, phenotype)

surv50_ris <- data.frame(strain = h_df_surv50$strain, 
                        phenotype = h_df_surv50$phenotype,
                        weibull_surv50_residual =residuals(lm(phenotype ~ block, data = h_df_surv50)))

h_df_surv90 <- h_df %>%
  dplyr::rename(phenotype = w_surv90_est)%>%
  dplyr::select(block, strain, phenotype)

surv90_ris <- data.frame(strain = h_df_surv90$strain, 
                         phenotype = h_df_surv90$phenotype,
                         weibull_surv90_residual =residuals(lm(phenotype ~ block, data = h_df_surv90)))

h_df_surv10 <- h_df %>%
  dplyr::rename(phenotype = w_surv10_est)%>%
  dplyr::select(block, strain, phenotype)

surv10_ris <- data.frame(strain = h_df_surv10$strain, 
                         phenotype = h_df_surv10$phenotype,
                         weibull_surv10_residual =residuals(lm(phenotype ~ block, data = h_df_surv10)))

h_df_corr <- as.data.frame(data.frame(strain = shape_ris[,1],
                   shape_par = shape_ris[,3],
                   scale_par = scale_ris[,3],
                   surv50_est = surv50_ris[,3],
                   sruv90_est = surv90_ris[,3],
                   surv10_est = surv10_ris[,3]))

h_df_corr <- h_df_corr %>%
  tidyr::gather(trait, phenotype, -strain) %>%
  as.data.frame()

glimpse(h_df_corr)
```

#### Calculate heritability for corrected traits
```{r}
# define heritability test
H2.test <- function(data){
  
  pheno <- as.data.frame(dplyr::select(data, phenotype))[,1]
  strain <- as.factor(data$strain)
  
  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))
  
  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))
  
  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)
  
  return(H2)
}

# apply heritability test to all block corrected traits
h_list <- list()

for(t in 1:length(unique(h_df_corr$trait))){
  
  trait_df <- h_df_corr %>%
    dplyr::filter(trait == (unique(h_df_corr$trait)[t])) %>%
    dplyr::mutate(H2 = H2.test(.))
  
  h_list[[t]] <- trait_df
}

h_cal_df <- dplyr::bind_rows(h_list) %>%
  dplyr::arrange(strain)

# plot heritability for each trait

herit_plot <- ggplot(h_cal_df%>%distinct(H2, .keep_all=T)) + 
  aes(x = trait, y = H2) +
  ylab("Heritability") +
  ylim(0, .5) +
  geom_point() +
  geom_hline(yintercept = .20, linetype="dashed", color="darkred")
herit_plot


# plot fit to one example
ggplot(df_proc1 %>% dplyr::filter(strain =="N2")) +
  aes(x=time_d, y=frac_alive, color=as.factor(bio_rep)) +
  geom_point() +
  theme(legend.position = "right") +
  stat_function(fun=function(x)exp(1)^-((x/13.11627)^3.763897), geom="line", color = "black") +
  theme_classic() +
  theme(title = element_text(size = 12, color = "black", face = "bold"),
        axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 12, color = "black"),
        legend.text = element_text(size = 12, color = "black")) + 
  labs(title = "", color = "Assay", x = "Day", y = "Fraction Alive") +
  theme(legend.position="right")

```
<br>
scale_par = weibull fit scale parameter<br>
shape_par = weibull fit shape parameter<br>
surv90_est = estimate for time(d) for 90% survival<br>
surv10_est = estimate for time(d) for 10% survival<br>
surv50_est = estimate for time(d) for 50% survival<br>