---
title: "Wild Isolate lifespan-healthspan"
author: "Tim C"
date: "05/05/2019"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r, echo=F, warning = F, message=F}
#library(tidyverse)
library(ggplot2)
library(dplyr)
library(viridis)
library(boot)
# set color palette
strain.colours <- c("DL238"="gold2", "N2"="plum4","JU775"= "darkorange1", 
                      "MY16"="lightskyblue2", "CX11314"="firebrick","ED3017"= "burlywood3", "LKC34"="gray51", 
                      "JU258"="springgreen4", "JT11398"="lightpink2", "CB4856"="deepskyblue4", 
                      "MY23"="mediumpurple4","EG4725"= "chocolate")

# load wormwatcher data
ww_df <- data.table::fread('~/Dropbox/AndersenLab/LabFolders/Tim/projects/WorMotel/data/20190430_wormwatcher_1.csv', header = T)

# load life machine data
lm_df <- data.table::fread('~/Dropbox/AndersenLab/LabFolders/Tim/projects/WorMotel/data/CeNDR_lifespans.csv', header = T)

# define heritability test
# data is data frame that contains strain and phenotype column
# indicies are used by the boot function to sample from the 'data' data.frame
H2.test.boot <- function(data, indicies){

  d <- data[indicies,]

  pheno <- as.data.frame(dplyr::select(d, phenotype))[,1]
  strain <- as.factor(d$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}
# data is data frame that contains strain and phenotype column
H2.test <- function(data){

  pheno <- as.data.frame(dplyr::select(data, phenotype))[,1]
  strain <- as.factor(data$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}

# df is data frame that contains strain and phenotype column
H2.calc <- function(df, boot = T){


  if(boot == T){
    # bootstrapping with 1000 replications
    results <- boot(data=df, statistic=H2.test.boot, R=500)

    # get 95% confidence interval
    ci <- boot.ci(results, type="bca")

    H2_errors <- data.frame(H2 = ci$t0, ci_l = ci$bca[4], ci_r = ci$bca[5])

    return(H2_errors)

  } else {

    H2 <- data.frame(H2 = H2.test(data = df), ci_l = NA, ci_r = NA)
    return(H2)
  }

}

# remove outliers function (tukey's fences)
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

```

### WormWatcher setup
<br> WormWatcher data consists of lifespan and healthspan traits measured in Chris Fang-Yen's lab, see `2019-04-23_WildIsolateUpdates.pptx` file for details on the experimental platform. In short, the data were collected with the wormwatcher robot at UPenn using 24-well plates. All six 24-well plates have identical layout of 12 strains with 2 replicates per strain. The plot below shows each of six experiments and the lifespans for each well are plotted according to the color scale on the right.<br>
```{r, echo=FALSE, warning=FALSE, message=F}
# data shaping
ww_df_proc <- ww_df %>%
  dplyr::rename( strain = Strain,
                 Plate_Num = `Plate Num`,
                 ls = `Lifespan est (days)`,
                 hs = `Healthspan est (days)`,
                 t = `Time (days)`,
                 act = `Activity (a.u.)`) %>%
  dplyr::distinct(Plate_Num, strain, Well_Num, .keep_all = T) %>%
  dplyr::filter(ls != "NaN") %>%
  #dplyr::mutate(indicies = seq(1, length(t), by = 1)) %>%
  dplyr::group_by(strain) %>%
  dplyr::mutate(reps = n(),
                median_ls = median(ls),
                median_hs = median(hs)) %>%
  dplyr::ungroup() %>%
  tidyr::gather(trait, phenotype, -Plate_Num, -Plate_Name, -Well_Num, -strain, -t, -act, -reps, -median_ls, -median_hs) %>%
  dplyr::arrange(median_hs)
  
# remove outliers in new df
ww_df_proc_outliers_removed <- ww_df_proc %>%
  dplyr::group_by(strain, trait) %>%
  dplyr::mutate(phenotype = remove_outliers(phenotype)) %>%
  dplyr::ungroup() %>%
  dplyr::filter(!is.na(phenotype))

# flag outliers in old data frame
ww_df_proc <- ww_df_proc %>%
  dplyr::group_by(strain, trait) %>%
  dplyr::mutate(outlier = (remove_outliers(phenotype)),
                outlier = ifelse(is.na(outlier), TRUE, FALSE)) %>%
  dplyr::ungroup()

# shaping data for H2 with bootstrapping 
ww_df_H2_ls <- ww_df_proc %>%
  dplyr::filter(trait == "ls") 
ww_df_H2_hs <- ww_df_proc %>%
  dplyr::filter(trait == "hs") 

ww_df_H2_ls_outliers_removed <- ww_df_proc_outliers_removed %>%
  dplyr::filter(trait == "ls") 
ww_df_H2_hs_outliers_removed <- ww_df_proc_outliers_removed %>%
  dplyr::filter(trait == "hs") 

# calculate H2 for both traits
ww_ls_h2 <- H2.calc(ww_df_H2_ls, boot = T)
ww_hs_h2 <- H2.calc(ww_df_H2_hs, boot = T)

ww_ls_h2_outliers_removed <- H2.calc(ww_df_H2_ls_outliers_removed, boot = T)
ww_hs_h2_outliers_removed <- H2.calc(ww_df_H2_hs_outliers_removed, boot = T)

# join H2 plots
ww_ls_h2_full <- bind_rows(ww_ls_h2, ww_ls_h2_outliers_removed) %>%
  dplyr::mutate(trait = as.factor(c("WW_LS","WW_LS no outliers")))

ww_hs_h2_full <- bind_rows(ww_hs_h2, ww_hs_h2_outliers_removed) %>%
  dplyr::mutate(trait = as.factor(c("WW_HS","WW_HS no outliers")))

# ploting data
ww_ls_h2_plot <- ggplot(ww_ls_h2_full) +
  geom_point(aes(x = trait, y = H2), size = 3) +
  geom_errorbar(aes(ymin=ci_l, ymax=ci_r, x = trait), colour="black", width=0.05) +
  labs(x = "", y = "Heritability") +
  ylim(0, 1) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ww_hs_h2_plot <- ggplot(ww_hs_h2_full) +
  geom_point(aes(x = trait, y = H2), size = 3) +
  geom_errorbar(aes(ymin=ci_l, ymax=ci_r, x = trait), colour="black", width=0.05) +
  labs(x = "", y = "Heritability") +
  ylim(0, 1) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# design plot data shaping
ww_design <- ww_df_proc %>%
  dplyr::mutate(row = ifelse(Well_Num %in% c(1:6), "A",
                             ifelse(Well_Num %in% c(7:12), "B",
                                    ifelse(Well_Num %in% c(13:18), "C", "D"))),
                column = ifelse(Well_Num %in% c(1,7,13,19), 1,
                                ifelse(Well_Num %in% c(2,8,14,20), 2,
                                       ifelse(Well_Num %in% c(3,9,15,21), 3,
                                              ifelse(Well_Num %in% c(4,10,16,22), 4,
                                                     ifelse(Well_Num %in% c(5,11,17,23), 5, 6))))))

# plot design
ww_design_plot <- ggplot(ww_design %>% dplyr::filter(trait == "ls")) +
 aes(x = column, y = row) +
    geom_tile(aes(fill = as.numeric(phenotype))) +
    geom_text(aes(label=strain), size = 2) +
    facet_wrap(~Plate_Name) +
    theme_bw() +
    scale_fill_viridis(name="") +
  labs(x="", y="")
ww_design_plot

```

### WormWatcher traits
```{r, echo=FALSE, warning=FALSE, fig.width= 7.5, fig.height=7.5}
median_ls_order <- c("DL238",   "N2", "JU775",
                     "MY16",    "CX11314", "ED3017",
                     "LKC34",   "JU258",   "JT11398",
                     "CB4856",  "EG4725", "MY23") 

median_hs_order <- c("JU775",   "DL238",   "N2",
                     "JU258",   "ED3017",  "JT11398",
                     "CX11314", "MY16",    "LKC34",
                     "CB4856",  "MY23",    "EG4725") 


# plot survival curves for all strains
ww_ls <- ggplot(ww_df_proc %>% dplyr::filter(trait == "ls")) +
  aes(x=factor(strain, levels = median_ls_order), y=phenotype, fill=strain) +
  scale_fill_manual(values = strain.colours) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="WW_Lifespan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ww_hs <- ggplot(ww_df_proc %>% dplyr::filter(trait == "hs")) +
  aes(x=factor(strain, levels = median_hs_order), y=phenotype, fill=strain) +
  scale_fill_manual(values = strain.colours) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="WW_Healthspan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

full_ww <- cowplot::plot_grid(ww_ls, ww_ls_h2_plot, ww_hs, ww_hs_h2_plot,
                          rel_widths = c(1, .25), ncol = 2, align = "vh", labels = c("A","", "B",""))
full_ww
```
<br> **(A-B)** Box plots for traits measured using WormWatcher (WW). **(A)** broad-sense heritability = 0.103 for Lifespan, but with outliers removed heritability = 0.257. **(B)** broad-sense heritability = 0.327 for Healthspan, with outlier removal 0.392.

### LifeMachine setup
These data are from Patrick Phillips lab. The data include individual worm lifespans for 20 strains assayed across 13 experiments. The plot below shows the randomized layout of the strains for the 13 experiments. The mean lifespans for individual agar plates are plotted according to the color scale on the right.
```{r, echo=FALSE, warning=FALSE, message=F, fig.width= 7.5, fig.height=7.5}
# what strains are in both data sets (All but DL238 is mis named PL238)
 similar <- lm_df %>%
  dplyr::filter(strain %in% names(strain.colours)) %>%
  dplyr::distinct(strain)

# How many censored
censored <- lm_df %>%
  dplyr::filter(censor == 1) %>%
  dplyr::group_by(strain) %>%
  dplyr::mutate(sum_censor = sum(censor)) %>%
  dplyr::distinct(strain, sum_censor)

# How many reps per experiment per strain
lm_design <- lm_df %>%
  dplyr::filter(censor != 1) %>%
  dplyr::group_by(exp, device, plate, strain) %>%
  dplyr::mutate(mean_ls = mean(ls),
                mean_DNMF = mean(DNMF)) %>%
  dplyr::ungroup()

lm_design_plot <- ggplot(lm_design %>% dplyr::distinct(exp,device, plate, strain, .keep_all = T)) +
 aes(x = column, y = row) +
    geom_tile(aes(fill = as.numeric(mean_ls))) +
    geom_text(aes(label=strain), size = 2) +
    facet_wrap(~exp) +
    theme_bw() +
  labs(x="", y="") +
    scale_fill_viridis(name="") 
lm_design_plot
```

### LifeMachine traits
```{r, echo=FALSE, warning=FALSE, fig.width= 7.5, fig.height=7.5}
###############
# data shaping
lm_df_proc <- lm_design %>%
  dplyr::distinct(exp,device, plate, strain, .keep_all = T) %>%
  dplyr::mutate(strain = ifelse(strain == "EC19191", "ECA191",
                                ifelse(strain == "PL238", "DL238", strain))) %>%
  dplyr::group_by(strain) %>%
  dplyr::mutate(strain_median = median(mean_ls)) %>%
  dplyr::arrange(strain_median)

median_ls_order_lm <- c("ECA191", "XZ1516",  "ECA744",  "ECA396",
"DL238",   "N2",      "JU258",   "MY16",
"ECA740",   "LKC34",  "JU2526",  "JU775",
"ED3017", "ECA36",   "EG4725",  "JT11398",
"CB4856", "JU2600",    "MY23",  "CX11314")


# Reshape data frame for H2 on 12 strain set
lm_ls_H2 <- lm_df_proc %>%
  dplyr::rename(phenotype = mean_ls) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(strain = factor(strain, levels = median_ls_order_lm)) %>%
  dplyr::mutate(trait = "ls")

# Same with CX11314 pruned
lm_ls_H2_prune <- lm_df_proc %>%
  dplyr::rename(phenotype = mean_ls) %>%
  dplyr::filter(strain != "CX11314") %>%
  dplyr::ungroup() %>%
  dplyr::mutate(strain = factor(strain, levels = median_ls_order_lm)) %>%
  dplyr::mutate(trait = "ls")

# calculate Heritability for both
lm_ls_h2 <- H2.calc(lm_ls_H2, boot = T)
lm_ls_h2_prune <- H2.calc(lm_ls_H2_prune, boot = T)

# Plot H2s
lm_ls_h2_plot <- ggplot(lm_ls_h2) +
  geom_point(aes(x = as.factor("LM Lifespan"), y = H2), size = 3) +
  geom_errorbar(aes(ymin=ci_l, ymax=ci_r, x = as.factor("LM Lifespan")), colour="black", width=0.05) +
  labs(x = "", y = "Heritability") +
  ylim(0, 1) +
  theme_bw()

lm_ls_h2_prune_plot <- ggplot(lm_ls_h2_prune) +
  geom_point(aes(x = as.factor("LM Lifespan"), y = H2), size = 3) +
  geom_errorbar(aes(ymin=ci_l, ymax=ci_r, x = as.factor("LM Lifespan")), colour="black", width=0.05) +
  labs(x = "", y = "Heritability") +
  ylim(0, 1) +
  theme_bw()

#plot box plots all
lm_ls <- ggplot(lm_df_proc %>% dplyr::filter(strain %in% names(strain.colours))) +
  aes(x=factor(strain, levels = median_ls_order_lm), y=mean_ls, fill=strain) +
  scale_fill_manual(values = strain.colours) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="LM_Lifespan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# with pruning
lm_df_proc_prune <- lm_df_proc %>%
  dplyr::filter(strain %in% names(strain.colours)) %>%
  dplyr::filter(strain != "CX11314")

lm_ls_prune <- ggplot(lm_df_proc_prune) +
  aes(x=factor(strain, levels = median_ls_order_lm), y=mean_ls, fill=strain) +
  scale_fill_manual(values = strain.colours) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="LM_Lifespan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
full_lm <- cowplot::plot_grid(lm_ls, lm_ls_h2_plot, lm_ls_prune, lm_ls_h2_prune_plot,
                          rel_widths = c(1, .25), ncol = 2, align = "vh", labels = c("A", "", "B", ""))

full_lm
```
<br> **(A)** Lifespan measured by life machine in Patrick Phillips lab, broad-sense heritability = 0.753. **(B)** The removal of CX11314 lowers broad-sense heritability to 0.674. Data points are mean lifespan for all individuals in plate. 

### Platform correlation
```{r, echo=FALSE, warning=FALSE, message=F, fig.width= 7.5, fig.height=3}
# astructure data for correlation tests

lm_corr <- lm_df_proc %>%
  dplyr::filter(strain %in% names(strain.colours)) %>%
  dplyr::distinct(strain, strain_median) %>%
  dplyr::rename(median_ls = strain_median) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(strain = factor(strain, levels = median_ls_order_lm)) %>%
  dplyr::mutate(ls_rank = seq(1, 12, 1)) %>%
  stats::setNames (paste0('lm_', names(.)))

ww_scorr <- ww_df_proc %>%
  dplyr::filter(trait == "ls") %>%
  dplyr::distinct(strain, median_ls) %>%
  dplyr::arrange(median_ls) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(strain = factor(strain, levels = median_ls_order)) %>%
  dplyr::mutate(ls_rank = seq(1, 12, 1)) %>%
  stats::setNames (paste0('ww_', names(.))) %>%
  dplyr::mutate(ww_strain = factor(ww_strain, levels = median_ls_order_lm)) %>%
  dplyr::arrange(ww_strain)

ww_pcorr <- ww_scorr %>%
  dplyr::mutate(ww_strain = factor(ww_strain, levels = median_ls_order_lm)) %>%
  dplyr::arrange(ww_strain) 

pcorr_df <- cbind(lm_corr, ww_pcorr)
scorr_df <- cbind(lm_corr, ww_scorr)

pcorr_df_prune <- pcorr_df %>% dplyr::filter(lm_strain != "CX11314")
scorr_df_prune <- scorr_df %>% dplyr::filter(lm_strain != "CX11314")

stest <- cor.test(x = scorr_df$lm_ls_rank, y = scorr_df$ww_ls_rank)
ptest <- cor.test(x = pcorr_df$lm_median_ls, y = pcorr_df$ww_median_ls)

stest_prune <- cor.test(x = scorr_df_prune$lm_ls_rank, y = scorr_df_prune$ww_ls_rank)
ptest_prune <- cor.test(x = pcorr_df_prune$lm_median_ls, y = pcorr_df_prune$ww_median_ls)

# plot correlation
p_plot <- ggplot(pcorr_df, aes(x = lm_median_ls, y = ww_median_ls)) +
  stat_smooth(method = 'lm', se = FALSE, color = "black", size = 0.5) +
  geom_point(aes(x = lm_median_ls, y = ww_median_ls, fill = lm_strain), shape = 21, size = 3) +
  scale_fill_manual(values = strain.colours) +
  labs(x="Lifespan LM", y="Lifespan WW") +
  theme_bw() +
  xlim(14,30) +
  ylim(14,30) +
  theme(legend.position = "none") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  # xlim(0,30) +
  # ylim(0,30) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #geom_text(x = 27.5, y = 15.25, label = paste0("r = ", round(as.numeric(ptest[4]), digits = 3)), size = 3, color = "black") +
  geom_text(x = 25.5, y = 15.25, label = paste0("ρ = ", round(as.numeric(stest[4]), digits = 3)), size = 3, color = "black")

# plot correlation with CX11314 removed
p_plot_rm <- ggplot(pcorr_df, aes(x = lm_median_ls, y = ww_median_ls)) +
  stat_smooth(data = pcorr_df %>% dplyr::filter(lm_strain != "CX11314"), aes(x = lm_median_ls, y = ww_median_ls), method = 'lm', se = FALSE, color = "red", size = 0.5) +
  geom_point(aes(x = lm_median_ls, y = ww_median_ls, fill = lm_strain), shape = 21, size = 3) +
  geom_point(aes(x = 29.60182, y = 17.10, color = "red"), shape = 4, size = 4) +
  scale_fill_manual(values = strain.colours) +
  labs(x="Lifespan LM", y="Lifespan WW") +
  theme_bw() +
  xlim(14,30) +
  ylim(14,30) +
  theme(legend.position = "none") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  # xlim(0,30) +
  # ylim(0,30) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #geom_text(x = 27.5, y = 15.25, label = paste0("r = ", round(as.numeric(ptest_prune[4]), digits = 3)), size = 3, color = "red") +
  geom_text(x = 25.5, y = 15.25, label = paste0("ρ = ", round(as.numeric(stest_prune[4]), digits = 3)), size = 3, color = "red")

full_corr <- cowplot::plot_grid(p_plot, p_plot_rm, ncol = 2, align = "vh", labels = c("A", "B"))
full_corr
```
<br> **(A-B)** Correlation of median lifespan estimates between platforms. **(A)** Spearman correlation coefficient with all strains included. **(B)** Same but with CX11314 removed from analysis.

### Heritability calc.
Fit linear mixed-effects model (LMM) to data with Strain as random variable.<br>
`reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))`<br>
<br>
Extract variance components for strain (Vg), and error (Ve).<br>
`Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))`<br>
`Vg <- Variances$vcov[1]`<br>
`Ve <- Variances$vcov[2]`<br>
<br>
Calculate broad-sense heritability with variance components.<br>
`H2 <- Vg/(Vg+Ve)`<br>
<br>
Calculate 95% confidence intervals by bootstrap method. 

```{r, echo=FALSE, warning=FALSE, message=F, fig.width= 7.5, fig.height=3}
# define heritability test
# data is data frame that contains strain and phenotype column
# indicies are used by the boot function to sample from the 'data' data.frame
H2.test.boot <- function(data, indicies){

  d <- data[indicies,]

  pheno <- as.data.frame(dplyr::select(d, phenotype))[,1]
  strain <- as.factor(d$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}
# data is data frame that contains strain and phenotype column
H2.test <- function(data){

  pheno <- as.data.frame(dplyr::select(data, phenotype))[,1]
  strain <- as.factor(data$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}

# df is data frame that contains strain and phenotype column
H2.calc <- function(df, boot = T){


  if(boot == T){
    # bootstrapping with 1000 replications
    results <- boot(data=df, statistic=H2.test.boot, R=500)

    # get 95% confidence interval
    ci <- boot.ci(results, type="bca")

    H2_errors <- data.frame(H2 = ci$t0, ci_l = ci$bca[4], ci_r = ci$bca[5])

    return(H2_errors)

  } else {

    H2 <- data.frame(H2 = H2.test(data = df), ci_l = NA, ci_r = NA)
    return(H2)
  }

}

# remove outliers function (tukey's fences)
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}
```
