---
title: "20190625_lifespan"
author: "Tim C."
date: "6/26/2019"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r, echo=F, warning = F, message=F}
#library(tidyverse)
library(ggplot2)
library(dplyr)
library(viridis)
library(boot)
# set color palette
strain.colours <- c("DL238"="gold2", "N2"="plum4","JU775"= "darkorange1", 
                      "MY16"="lightskyblue2", "CX11314"="firebrick","ED3017"= "burlywood3", "LKC34"="gray51", 
                      "JU258"="springgreen4", "JT11398"="lightpink2", "CB4856"="deepskyblue4", 
                      "MY23"="mediumpurple4","EG4725"= "chocolate")

# load wormwatcher data
ww_df <- data.table::fread('~/Dropbox/AndersenLab/LabFolders/Tim/projects/WorMotel/data/2019-05-24_AllDataLongFormat.csv', header = T)

# load life machine data
lm_df <- data.table::fread('~/Dropbox/AndersenLab/LabFolders/Tim/projects/WorMotel/data/CeNDR_lifespans.csv', header = T)

# define heritability test
# data is data frame that contains strain and phenotype column
# indicies are used by the boot function to sample from the 'data' data.frame
H2.test.boot <- function(data, indicies){

  d <- data[indicies,]

  pheno <- as.data.frame(dplyr::select(d, phenotype))[,1]
  strain <- as.factor(d$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}
# data is data frame that contains strain and phenotype column
H2.test <- function(data){

  pheno <- as.data.frame(dplyr::select(data, phenotype))[,1]
  strain <- as.factor(data$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}

# df is data frame that contains strain and phenotype column
H2.calc <- function(df, boot = T){


  if(boot == T){
    # bootstrapping with 1000 replications
    results <- boot(data=df, statistic=H2.test.boot, R=500)

    # get 95% confidence interval
    ci <- boot.ci(results, type="bca")

    H2_errors <- data.frame(H2 = ci$t0, ci_l = ci$bca[4], ci_r = ci$bca[5])

    return(H2_errors)

  } else {

    H2 <- data.frame(H2 = H2.test(data = df), ci_l = NA, ci_r = NA)
    return(H2)
  }

}

# remove outliers function (tukey's fences)
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

```

### WormWatcher setup
<br> WormWatcher data consists of lifespan and healthspan traits measured in Chris Fang-Yen's lab, see `2019-05-24_WildIsolateUpdates.pptx` file for details on the experimental platform. In short, the data were collected with the wormwatcher robot at UPenn using 24-well plates. All 14 24-well plates have 12 strains with 2 replicates per strain. The plot below shows each of 14 plates and the lifespans for each well are plotted according to the color scale on the right. I am not breaking analysis out by unique bleach, but we could do this in the future.<br>
```{r, echo=FALSE, warning=FALSE, message=F}
# data shaping
ww_df_proc <- ww_df %>%
  dplyr::rename( strain = Strain,
                 Plate_Num = `Plate Num`,
                 ls = `Lifespan est (days)`,
                 hs = `Healthspan est (days)`,
                 t = `Time (days)`,
                 act = `Activity (a.u.)`) %>%
  dplyr::distinct(Plate_Num, Plate_Name, strain, Well_Num, .keep_all = T) %>%
  dplyr::filter(ls != "NaN") %>%
  #dplyr::mutate(indicies = seq(1, length(t), by = 1)) %>%
  dplyr::group_by(strain) %>%
  dplyr::mutate(reps = n(),
                median_ls = median(ls),
                median_hs = median(hs)) %>%
  dplyr::ungroup() %>%
  tidyr::gather(trait, phenotype, -Plate_Num, -Plate_Name, -Well_Num, -strain, -t, -act, -reps, -median_ls, -median_hs) %>%
  dplyr::arrange(median_hs)
  
# remove outliers in new df
ww_df_proc_outliers_removed <- ww_df_proc %>%
  dplyr::group_by(strain, trait) %>%
  dplyr::mutate(phenotype = remove_outliers(phenotype)) %>%
  dplyr::ungroup() %>%
  dplyr::filter(!is.na(phenotype))

# flag outliers in old data frame
ww_df_proc <- ww_df_proc %>%
  dplyr::group_by(strain, trait) %>%
  dplyr::mutate(outlier = (remove_outliers(phenotype)),
                outlier = ifelse(is.na(outlier), TRUE, FALSE)) %>%
  dplyr::ungroup()

# shaping data for H2 with bootstrapping 
ww_df_H2_ls <- ww_df_proc %>%
  dplyr::filter(trait == "ls") 
ww_df_H2_hs <- ww_df_proc %>%
  dplyr::filter(trait == "hs") 

ww_df_H2_ls_outliers_removed <- ww_df_proc_outliers_removed %>%
  dplyr::filter(trait == "ls") 
ww_df_H2_hs_outliers_removed <- ww_df_proc_outliers_removed %>%
  dplyr::filter(trait == "hs") 

# calculate H2 for both traits
ww_ls_h2 <- H2.calc(ww_df_H2_ls, boot = T)
ww_hs_h2 <- H2.calc(ww_df_H2_hs, boot = T)

ww_ls_h2_outliers_removed <- H2.calc(ww_df_H2_ls_outliers_removed, boot = T)
ww_hs_h2_outliers_removed <- H2.calc(ww_df_H2_hs_outliers_removed, boot = T)

# join H2 plots
ww_ls_h2_full <- bind_rows(ww_ls_h2, ww_ls_h2_outliers_removed) %>%
  dplyr::mutate(trait = as.factor(c("WW_LS","WW_LS no outliers")))

ww_hs_h2_full <- bind_rows(ww_hs_h2, ww_hs_h2_outliers_removed) %>%
  dplyr::mutate(trait = as.factor(c("WW_HS","WW_HS no outliers")))

# ploting data
ww_ls_h2_plot <- ggplot(ww_ls_h2_full) +
  geom_point(aes(x = trait, y = H2), size = 3) +
  geom_errorbar(aes(ymin=ci_l, ymax=ci_r, x = trait), colour="black", width=0.05) +
  labs(x = "", y = "Heritability") +
  ylim(0, 1) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ww_hs_h2_plot <- ggplot(ww_hs_h2_full) +
  geom_point(aes(x = trait, y = H2), size = 3) +
  geom_errorbar(aes(ymin=ci_l, ymax=ci_r, x = trait), colour="black", width=0.05) +
  labs(x = "", y = "Heritability") +
  ylim(0, 1) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# design plot data shaping
ww_design <- ww_df_proc %>%
  dplyr::mutate(row = ifelse(Well_Num %in% c(1:6), "A",
                             ifelse(Well_Num %in% c(7:12), "B",
                                    ifelse(Well_Num %in% c(13:18), "C", "D"))),
                column = ifelse(Well_Num %in% c(1,7,13,19), 1,
                                ifelse(Well_Num %in% c(2,8,14,20), 2,
                                       ifelse(Well_Num %in% c(3,9,15,21), 3,
                                              ifelse(Well_Num %in% c(4,10,16,22), 4,
                                                     ifelse(Well_Num %in% c(5,11,17,23), 5, 6))))))

# plot design
ww_design_plot <- ggplot(ww_design %>% dplyr::filter(trait == "ls")) +
 aes(x = column, y = row) +
    geom_tile(aes(fill = as.numeric(phenotype))) +
    geom_text(aes(label=strain), size = 2) +
    facet_wrap(~Plate_Name) +
    theme_bw() +
    scale_fill_viridis(name="") +
  labs(x="", y="")
ww_design_plot

```


### WormWatcher traits
```{r, echo=FALSE, warning=FALSE, fig.width= 7.5, fig.height=7.5}
median_ls_ranks <- ww_df_proc %>%
  dplyr::distinct(strain, .keep_all = T) %>%
  dplyr::arrange(median_ls) %>%
  dplyr::mutate(ls_rank = seq(1:12)) 

median_hs_ranks <- ww_df_proc %>%
  dplyr::distinct(strain, .keep_all = T) %>%
  dplyr::arrange(median_hs) %>%
  dplyr::mutate(hs_rank = seq(1:12))

median_ls_order <- median_ls_ranks$strain

median_hs_order <- median_hs_ranks$strain

# median_ls_order <- c("DL238",   "N2", "JU775",
#                      "MY16",    "CX11314", "ED3017",
#                      "LKC34",   "JU258",   "JT11398",
#                      "CB4856",  "EG4725", "MY23") 
# 
# median_hs_order <- c("JU775",   "DL238",   "N2",
#                      "JU258",   "ED3017",  "JT11398",
#                      "CX11314", "MY16",    "LKC34",
#                      "CB4856",  "MY23",    "EG4725") 


# plot survival curves for all strains
ww_ls <- ggplot(ww_df_proc %>% dplyr::filter(trait == "ls")) +
  aes(x=factor(strain, levels = median_ls_order), y=phenotype, fill=strain) +
  scale_fill_manual(values = strain.colours) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="WW_Lifespan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ww_hs <- ggplot(ww_df_proc %>% dplyr::filter(trait == "hs")) +
  aes(x=factor(strain, levels = median_hs_order), y=phenotype, fill=strain) +
  scale_fill_manual(values = strain.colours) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="WW_Healthspan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

full_ww <- cowplot::plot_grid(ww_ls, ww_ls_h2_plot, ww_hs, ww_hs_h2_plot,
                          rel_widths = c(1, .25), ncol = 2, align = "vh", labels = c("A","", "B",""))
full_ww
```
<br> **(A-B)** Box plots for traits measured using WormWatcher (WW). Each data point is a well. **(A)** broad-sense heritability = 0.342 for Lifespan, removing outliers doesn't change heritability estimate just reduces confidence interval. **(B)** broad-sense heritability = 0.322 for Healthspan, with outlier removal 0.332.



### Heritability calc.
Fit linear mixed-effects model (LMM) to data with Strain as random variable.<br>
`reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))`<br>
<br>
Extract variance components for strain (Vg), and error (Ve).<br>
`Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))`<br>
`Vg <- Variances$vcov[1]`<br>
`Ve <- Variances$vcov[2]`<br>
<br>
Calculate broad-sense heritability with variance components.<br>
`H2 <- Vg/(Vg+Ve)`<br>
<br>
Calculate 95% confidence intervals by bootstrap method. 

### WormWatcher traits for plate and bleach averages
<br> There are seven independent bleaches in the large data set, four conducted by JX, three by JDH. To get the mean value per bleach per strain I average the data for replicate wells in a plate (2 wells), then average those across technical replicate plates (2 plates). I've dropped one bleach from JX b/c many wells were contaminated. This gives us five to six independent bleaches per strain for analysis.<br>
```{r, echo=FALSE, warning=FALSE, message=F}
# data shaping
ww_df_proc_2 <- ww_df %>%
  dplyr::rename( strain = Strain,
                 Plate_Num = `Plate Num`,
                 ls = `Lifespan est (days)`,
                 hs = `Healthspan est (days)`,
                 t = `Time (days)`,
                 act = `Activity (a.u.)`) %>%
  dplyr::select(-act, -t) %>%
  dplyr::distinct(Plate_Num, Plate_Name, strain, Well_Num, .keep_all = T) %>%
  dplyr::filter(ls != "NaN") %>%
  #dplyr::mutate(indicies = seq(1, length(t), by = 1)) %>%
  # dplyr::group_by(strain) %>%
  # dplyr::mutate(reps = n(),
  #               median_ls = median(ls),
  #               median_hs = median(hs)) %>%
  # dplyr::ungroup() %>%
  # tidyr::gather(trait, phenotype, -Plate_Num, -Plate_Name, -Well_Num, -strain, -t, -act, -reps, -median_ls, -median_hs) %>%
  # dplyr::arrange(median_hs) %>%
  tidyr::separate(Plate_Name, c("blah2", "plate_num2", "initials2"), "_") %>%
  dplyr::mutate(plate = ifelse(Plate_Num >= 9, initials2, plate_num2),
                initials = ifelse(Plate_Num >= 9, plate_num2, initials2)) %>%
  dplyr::select(-blah2, -plate_num2, -initials2) %>%
  dplyr::mutate(bleach = ifelse(grepl("6",plate), 6,
                                    ifelse(grepl("7",plate), 7,
                                           ifelse(grepl("8",plate), 8,
                                                  ifelse(grepl("9",plate), 9, NA))))) %>%
  dplyr::mutate(tech_rep = ifelse(grepl("A",plate), "A",
                                    ifelse(grepl("B",plate), "B", NA))) %>%
  dplyr::select(-plate)
  
#remove bleach 6 b/c its only 1 prep and many wells are contaminated and average across plates and bleaches
ww_df_proc_3 <- ww_df_proc_2 %>%
  dplyr::filter(bleach != 6) %>%
  dplyr::group_by(strain, Plate_Num) %>%
  dplyr::mutate(strain_plate_mean_ls = mean(ls),
                strain_plate_mean_hs = mean(hs)) %>%
  dplyr::group_by(strain, initials, bleach) %>%
  dplyr::mutate(strain_initials_bleach_mean_ls = mean(unique(strain_plate_mean_ls)),
                strain_initials_bleach_mean_hs = mean(unique(strain_plate_mean_hs))) %>%
  dplyr::group_by(strain, bleach) %>%
  dplyr::mutate(strain_bleach_mean_ls = mean(unique(strain_initials_bleach_mean_ls)),
                strain_bleach_mean_hs = mean(unique(strain_initials_bleach_mean_hs))) %>%
  dplyr::ungroup()

# setup data with plate means and bleach means
ww_df_proc_4 <- ww_df_proc_3 %>%
  dplyr::distinct(strain, strain_initials_bleach_mean_ls, .keep_all = T) %>%
  dplyr::select(-strain_plate_mean_ls, -strain_plate_mean_hs, -strain_bleach_mean_ls, -strain_bleach_mean_hs, -Well_Num, -Plate_Num, -ls, -hs, -tech_rep) %>%
  tidyr::gather(trait, phenotype, -strain, -initials, -bleach)

# # remove outliers in new df
# ww_df_proc_outliers_removed <- ww_df_proc %>%
#   dplyr::group_by(strain, trait) %>%
#   dplyr::mutate(phenotype = remove_outliers(phenotype)) %>%
#   dplyr::ungroup() %>%
#   dplyr::filter(!is.na(phenotype))
# 
# # flag outliers in old data frame
# ww_df_proc <- ww_df_proc %>%
#   dplyr::group_by(strain, trait) %>%
#   dplyr::mutate(outlier = (remove_outliers(phenotype)),
#                 outlier = ifelse(is.na(outlier), TRUE, FALSE)) %>%
#   dplyr::ungroup()

# shaping data for H2 with bootstrapping 
ww_df_H2_ls_pbmeans <- ww_df_proc_4 %>%
  dplyr::filter(trait == "strain_initials_bleach_mean_ls") %>%
  dplyr::select(-initials, -bleach)
ww_df_H2_hs_pbmeans <- ww_df_proc_4 %>%
  dplyr::filter(trait == "strain_initials_bleach_mean_hs") %>%
  dplyr::select(-initials, -bleach)

# ww_df_H2_ls_outliers_removed <- ww_df_proc_outliers_removed %>%
#   dplyr::filter(trait == "ls") 
# ww_df_H2_hs_outliers_removed <- ww_df_proc_outliers_removed %>%
#   dplyr::filter(trait == "hs") 

# calculate H2 for both traits
ww_ls_h2_pbmeans <- H2.calc(ww_df_H2_ls_pbmeans, boot = T)
ww_hs_h2_pbmeans <- H2.calc(ww_df_H2_hs_pbmeans, boot = T)

# ww_ls_h2_outliers_removed <- H2.calc(ww_df_H2_ls_outliers_removed, boot = T)
# ww_hs_h2_outliers_removed <- H2.calc(ww_df_H2_hs_outliers_removed, boot = T)

# join H2 plots
ww_ls_h2_full_pbmeans <- ww_ls_h2_pbmeans %>%
  dplyr::mutate(trait = as.factor(c("Life_span")))

ww_hs_h2_full_pbmeans <- ww_hs_h2_pbmeans %>%
  dplyr::mutate(trait = as.factor(c("Health_span")))

full_pbmeans_hsls <- full_join(ww_ls_h2_full_pbmeans, ww_hs_h2_full_pbmeans)

# ploting heritability data
ww_ls_h2_plot <- ggplot(full_pbmeans_hsls) +
  geom_point(aes(x = trait, y = H2), size = 3) +
  geom_errorbar(aes(ymin=ci_l, ymax=ci_r, x = trait), colour="black", width=0.05) +
  labs(x = "", y = "Broad-sense Heritability") +
  ylim(0, 1) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept = .2, color= "dark red", lty = 2)

ls_order = c("N2", "ED3017","CX11314","JU775","DL238",
             "LKC34","MY16","JU258","MY23","JT11398",
             "CB4856","EG4725")

hs_order = c("N2", "JU775","DL238","JU258","ED3017","CX11314",
             "LKC34","JT11398","MY16","CB4856","MY23","EG4725")

# plotting assay data
ww_ls <- ggplot(ww_df_proc_4 %>% dplyr::filter(trait == "strain_initials_bleach_mean_ls")) +
  aes(x=factor(strain, levels = ls_order), y=phenotype, fill=strain) +
  scale_fill_manual(values = strain.colours) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="WW_Lifespan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ww_hs <- ggplot(ww_df_proc_4 %>% dplyr::filter(trait == "strain_initials_bleach_mean_hs")) +
  aes(x=factor(strain, levels = hs_order), y=phenotype, fill=strain) +
  scale_fill_manual(values = strain.colours) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="WW_Healthspan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

full_ww <- cowplot::plot_grid(ww_ls,  ww_ls_h2_plot, ww_hs,
                          rel_widths = c(1, .25), ncol = 2, align = "vh", labels = c("A", "C","B"))
full_ww

# plot bleach effects
bleach_ls <- ggplot(ww_df_proc_4 %>% dplyr::filter(trait == "strain_initials_bleach_mean_ls") %>%
                      tidyr::unite(bleach, c("bleach", "initials"))) +
  aes(x = as.factor(bleach), y = phenotype) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = strain.colours) +
  geom_jitter(width = .25, shape = 21) +
  labs(x="", y="WW_Healthspan") +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# # setup interaction plot  
# ww_df_proc_5_ls <- ww_df_proc_4 %>%
#   dplyr::filter(trait == "strain_initials_bleach_mean_ls") %>%
#                       tidyr::unite(bleach, c("bleach", "initials"))
# 
# test <- interaction.plot(x.factor = ww_df_proc_5_ls$bleach,    # variable to plot on x-axis
#                  trace.factor = ww_df_proc_5_ls$strain, # variable to specify "traces"; here, lines
#                  response = ww_df_proc_5_ls$phenotype,    # variable to plot on y-axis
#                  fun = median,  # summary statistic to be plotted for response variable
#                  type = "l",     # type of plot, here "l" for lines
#                  ylab = "Life_span",
#                  xlab = "Independent bleaches",
#                  col = strain.colours,
#                  lty = 1,  # line type
#                  lwd = 2,  # line width
#                  trace.label = "strain",  # label for legend
#                  xpd = T) #,  # 'clip' legend at border

```

```{r, echo=FALSE, warning=FALSE, message=F, fig.width= 7.5, fig.height=3}
# define heritability test
# data is data frame that contains strain and phenotype column
# indicies are used by the boot function to sample from the 'data' data.frame
H2.test.boot <- function(data, indicies){

  d <- data[indicies,]

  pheno <- as.data.frame(dplyr::select(d, phenotype))[,1]
  strain <- as.factor(d$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}
# data is data frame that contains strain and phenotype column
H2.test <- function(data){

  pheno <- as.data.frame(dplyr::select(data, phenotype))[,1]
  strain <- as.factor(data$strain)

  reffMod <- lme4::lmer(pheno ~ 1 + (1|strain))

  Variances <- as.data.frame(lme4::VarCorr(reffMod, comp = "Variance"))

  Vg <- Variances$vcov[1]
  Ve <- Variances$vcov[2]
  H2 <- Vg/(Vg+Ve)

  # errors <- sqrt(diag(lme4::VarCorr(reffMod, comp = "Variance")$strain))

  return(H2)
}

# df is data frame that contains strain and phenotype column
H2.calc <- function(df, boot = T){


  if(boot == T){
    # bootstrapping with 1000 replications
    results <- boot(data=df, statistic=H2.test.boot, R=500)

    # get 95% confidence interval
    ci <- boot.ci(results, type="bca")

    H2_errors <- data.frame(H2 = ci$t0, ci_l = ci$bca[4], ci_r = ci$bca[5])

    return(H2_errors)

  } else {

    H2 <- data.frame(H2 = H2.test(data = df), ci_l = NA, ci_r = NA)
    return(H2)
  }

}

# remove outliers function (tukey's fences)
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}
```
